### 2014-01-26 foo.tex
more tutorial note

\documentclass[12pt,letterpaper]{article}
\usepackage{listings}
\lstset{
frame = single,
language = bash,
breaklines = true,
postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\color{red}\hookrightarrow\space}}
}

\title{Git Learning Notes}
\author{tao}
\date{\today}

\begin{document}

\maketitle
\section{New tracking project}
Minimum working sample is:
\begin{lstlisting}
Git init
Git add */*.tex
Git add *.tex
Git add readme
Git commit -m 'comments'
% on a specific file after editing
 git add xx.tex

% to Reset a staged file
Git reset HEAD xxx.tex

ssh-keygen -t rsa -C "qisaiman@gmail.com"

% checkout from github 
git add remote hw https://github.com/qisaiman/hw.git
git fetch hw master or develop 
git push hw master
\end{lstlisting}

Git protocol 9148 port.

\section{Commands}

\begin{table}
  \caption{Commands}
  \label{tbl:notes}
  \begin{tabular}{ll}
    \toprule
    Command                            & Function \\
    \midrule
    gitk\textsuperscript{\emph{a}}   & Visualization versioning history  \\
    git status\textsuperscript{\emph{b}} & current working flow state \\
    ssh key   & zgcss627\\

    \bottomrule
  \end{tabular}

  \textsuperscript{\emph{a}} Some text;
  \textsuperscript{\emph{b}} Some more text.
\end{table}

\end{document}