### 2014-02-23 BackMatter/app3.tex
process applendix added

\chapter{Processes developed during Ph.D. study}

\section{Growth recipes}

\section{Characterization and data processing}

\subsection{UV-Vis measurements}

\subsection{TEM-Raman integrated measurements}

\subsection{data processing}

\textbf{Multipeaks fitting}

\textbf{3D pdf}