### 2014-09-27 ch-tms/tms.tex

\chapter{Tungsten Oxide-Tungsten Disulfide Heterostructure}

In this chapter, a TEM-Raman integrated study on \ce{WO3}-\ce{WS2} core-shell nanostructures is presented. This study leverages a home-built micromanipulator apparatus to transfer the nanostructures between TEM grids and other desired substrates. This technique has enabled us to correlate TEM and micro-Raman measurement on individual nanowires, thereby enhancing the understanding of structure-property relations which would otherwise be blurred in conventional ensemble averaged methods. \ce{WO3}-\ce{WS2} heterostructure is chosen due to the potential applications in photocatalysis and hydrogen evolution.

The remaining section is arranged in following order: a review on the growth of \ce{MS2} (M = Mo, W) nanotubes (NTs) and 2D \gls{fl} structures will be introduced first, followed by Raman spectroscopy summary on \ce{WS2} NTs and FL. The thesis will then focus on synthesis of \ce{WO3}-\ce{WS2} heterostructure and various characterizations on both ensemble and individual nanowires. This chapter is concluded with a brief summary and future work.

\input{./ch-tms/tmspart}

\section{Experimental}

Four kinds of tungsten powders were used as precursor to prepare \ce{WO3} NWs, and sulfur powder was used to fabricate \ce{WO3}-\ce{WS2} core-shell NWs, as summarized in Table.~\ref{tab:ch5pre}. All reactants were employed as received without further processing. Substrates cleaning procedures have been introduced in Sec.~\ref{ch2sub}.

% Tungsten powders size and purity
\begin{table}[htb]
\centering
\caption{List of reactants for \ce{WO3}-\ce{WS2}}\label{tab:ch5pre}
\begin{tabular}{lccr}
\toprule
Name & purity & average size & vendor info\\
\midrule
3N   &  99.9\% & 17 $\mu$m & Alfa Aesar \#39749\\
3N5   &  99.95\% & 32 $\mu$m  & Alfa Aesar \#42477\\
4N5   &  99.995\% & 3.3 $\mu$m  & Materion T-2049 \\
5N   &  99.999\% & 1.5 $\mu$m & Alfa Aesar \#12973\\
S    &   99.5\%  &  NA  & Alfa Aesar \#10785\\
\bottomrule
\end{tabular}
\end{table}

\ce{WO3} NWs were obtained by a seeded growth with tungsten powders applied onto the Si substrate. To find out the optimal conditions, tungsten powder oxidation experimental were carried out. Schematic layout of each growth were illustated in Fig.~\ref{fig:ch5grow} including the sulfurization process.

% cvd NW growth
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{cs_sfig1_all}
\caption[\ce{WO3} NW growth: OT]{Schematic diagrams of (a) tungsten power oxidization, (b) seeded growth of \ce{WO3} NW, and (c) sulfurization process. (d) Chemical vapor deposition system temperature profile measured at ambient condition. The zero inch location is defined at the upstream edge of furnace.}
\label{fig:ch5grow}
\end{figure}

In a typical oxidation experiment shown in Fig.~\ref{fig:ch5grow}a, tungsten powders were loaded into the uniform heating zone and the sealed chamber was pumped down to an ultimate pressure of $5\sim8$ mTorr. Then oxygen flow varying between 1 sccm to 10 sccm (standard cubic centimeter per minute) was admitted from upstream inlet. With 10 sccm UHP Ar (99.999\%) as carrier gas, the overall pressure reached to about 100 mTorr. The heating temperature (500 to 750 \si{\degreeCelsius}) was ramped up in 30 min and lasted for 30 min. Then the heating power was turned off and the chamber was allowed to naturally cool down to room temperature.

In a representative seeded growth (Fig.~\ref{fig:ch5grow}b), 4N5 or 5N tungsten source powder (Table.~\ref{tab:ch5pre}) was positioned in the upstream end of a quartz boat and downstream about 2.5 in the substrate was stationed. Additional tungsten seed powders (3N) were distributed evenly onto the substrate. After pumping down, 1 sccm oxygen and 10 sccm Ar were admitted into the chamber, respectively. The heating temperature was ramped up to 1000 \si{\degreeCelsius} in 30 min and lasted for 240 min. Subsequently the apparatus was allowed to naturally cool down to room temperature.

\ce{WO3}-\ce{WS2} core-shell NWs were synthesized using direct sulfurization of \ce{WO3} NWs. As depicted in Fig.~\ref{fig:ch5grow}c, the as-grown \ce{WO3} NWs were loaded into the center of heating furnace, and $\sim200$ mg sulfur (Alfa Aesar 10785, 99.5\%) was positioned just outside the upstream edge of furnace, where the maximum temperature was about 240 \si{\degreeCelsius}. After pumping down, the reaction chamber was flushed two times to expel residual air. A cold trap filled with liquid nitrogen in downstream was used to collect possible sulfur precipitation. Then the furnace was heated to 750 \si{\degreeCelsius} in 30 min, held for 15 min, and allowed to naturally cool down to room temperature. During entire growth process, 30 sccm Ar was used as carrier flow.

\section{Results and Discussion}
\subsection{Ensemble Measurements on Core-Shell NWs Array}

The morphology of \ce{WO3} NWs in micron scale almost stay the same after sulfurization, as shown in the inset SEM imaging of Fig.~\ref{fig:ch5ws2sem}. The length of individual nanowire can be up to 15 $\mu$m and the diameter varies from about 40 to 300 nm. The surface of the NWs were rather smooth without other attachments. The NWs array grow slightly larger than the W seed ( $\sim 12 \mu$m), mostly due to the elongation of NW.

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{ws2_sem_edx}
\caption[Morphology and composition analysis on \ce{WO3}-\ce{WS2} NWs]{Morphology and composition analysis on the as-synthesized core-shell structures. (a) SEM images showing the NWs almost retain their original shapes after sulfurization, and (b) EDX spectroscopy revealing the presence of sulfur element after sulfurization.}
\label{fig:ch5ws2sem}
\end{figure}

The EDX spectrum in Fig.~\ref{fig:ch5ws2sem} revealed the presence of sulfur element, indicating the effectiveness of sulfurization. Compared to the XRD pattern on seeded growth \ce{WO3} sample, post-sulfurization pattern (Fig.~\ref{fig:ch5ws2xrd}) shows a much reduced \ce{WO3} intensity with W phase becoming relatively prominent. Still the peaks at 23.44$^{\circ}$ and 24.36$^{\circ}$ can be indexed to \ce{WO3} (020) and (200) reflection, respectively. This result as well as the SEM imaging indicates \ce{WO3} NWs could preserve some long range order after the sulfurization for 15 min at 750 \si{\degreeCelsius}.

%fig xrd ws2 
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{ws2_cs_xrd}
\caption[XRD on core-shell NWs]{X-ray diffraction spectra before and after sulfurization. The reference spectra for $c$-W, 2H-\ce{WS2} and $m$-\ce{WO3} are included with some major peaks labeled.}
\label{fig:ch5ws2xrd}
\end{figure}

This observation can also be used to account for the extreme small peak at 13.87$^{\circ}$ in the post-sulfurization spectrum. Compared the reference 2H-\ce{WS2} (002) peak (14.37$^{\circ}$, ICDD PDF 04-003-4478), this redshift indicates an increase of lattice spacing along $c$ axis (from 6.16 \AA to 6.38 \AA). The FWHM of this (002) peak is about 1.1$^{\circ}$, giving an estimation on the crystalline size of 7.5 nm according to Scherrer equation.

Due to this small spacing, TEM is primarily utilized to observe the detailed morphology of \ce{WO3}-\ce{WS2} core-shell NWs. Fig.~\ref{fig:ch5ws2tem1} shows HRTEM imaging on typical tip and body region of the core-shell NW, respectively. A comparison with the pre-sulfurization specimen (Fig.~\ref{fig:woseedtem2}) readily reveals the formation of core-shell morphology after sulfurization. The measured layer spacing is 6.4 \AA, slight larger than bulk 2H-\ce{WS2} of 6.16 \AA, but in good agreement with XRD results in Fig.~\ref{fig:ch5ws2xrd}. The clear contrast in core region also prove the lattice structure of \ce{WO3} is not seriously distorted.

% tem core shell layer
\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{ws2_coreshell_tem}
\caption[TEM imaging on core-shell NWs: 1]{HRTEM imaging on one core-shell nanowire showing layer number changes.}
\label{fig:ch5ws2tem1}
\end{figure}

After examining dozens of core-shell NWs, this thesis found the \ce{WS2} layer number at NW tip was usually the largest, reaching more than ten, and the layer number could decrease to one along the NW body, but there exist some layer number fluctuation and shape modification. Fig.~\ref{fig:ch5ws2tem3} shows a smooth transition of \ce{WS2} layer number on the tip region of one NW. The layer number is about 7 on the tip corner, gradually decreases from 4 to 3 along the NW body, and stops at 2. Closer observation indicates the layer spacing is not quite uniform in the tip area, probably due to the complex curvature and bending. \ce{WS2} fullerene also show similar structures. Since the closing of the tip region could be of great complexity and the induced strain could modify Raman signal considerably, the TEM-Raman integrated measurement in this study will focus on the body region only, with those having smooth transition of layer number.
% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{ws2_growth3}
\caption[TEM imaging on core-shell NWs: 2]{HRTEM imaging on the tip area of one core-shell nanowire.}
\label{fig:ch5ws2tem3}
\end{figure}

As illustrated in Fig.~\ref{fig:ch5ws2tem4}, complicated \ce{WS2} encapsulation was observed on one tapered core-shell NW. This tapering probably arose from the sulfurization process, but could be also due to the parent \ce{WO3} NW. The diameter varied from about 200 nm to 15 nm, and the \ce{WS2} layer number also changed in a nonlinear manner. On some turning region, the \ce{WS2} seems penetrating into the body of primary \ce{WO3} NW.

% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{ws2_growth4}
\caption[TEM imaging on core-shell NWs: 3]{HRTEM imaging on one tapered core-shell nanowire showing fluctuation of layer numbers}
\label{fig:ch5ws2tem4}
\end{figure}

A systematic study on the sulfurization must be performed to understand the growth dynamics. The growth mechanism of \ce{WS2} nanotubes from oxides were thoroughly studied,\cite{Feldman1998} and it is generally agreed that the sulfurization of \ce{WO3} is a diffusion limited reaction.\cite{Feldman1996} It was also suggested that the core \ce{WO3} NWs will become non-stoichiometric phases upon sulfurization.\cite{Feldman1996,ZAK2009} This study recognizes this possibility but also consider two factors different from previous studies: one is the usage of sulfur as precursor instead of \ce{H2} and/or \ce{H2S} as in previous reports, which will accelerate the reduction of \ce{WO3} significantly due to the formation of \ce{H2O}; the other is the short reaction time and light degree of sulfurization, which probably will not distort the \ce{WO3} lattice significantly, as evidenced by the XRD and HRTEM analysis in this section. The sulfurization dynamic is then proposed as following in this thesis: the tips of \ce{WO3} NWs exhibit greater crystalline disorder, and \ce{WS2} layers on tips are folded at least along two axes, resulting a higher density of defects and thus a higher sulfurization rate; while the folding along NW body is quasi-one dimensional, subsequently the sulfurization rate could be lower. Besides, sulfurization could also initiated from the NW body, presumably due to the presence of surface defects. As a result, this led to the uneven \ce{WS2} layer number distribution along the core-shell NWs.

No obviously consumed core region was observed among most of the core-shell NWs examined under TEM. An exception is shown in Fig.~\ref{fig:ch5ws2tem2}. It is worth mentioning that some interesting growth behavior around the void region, where a new \ce{WS2} layer pushing through the already formed \ce{WS2} ones, deviating from the diffusion growth model.

% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{ws2_growth1}
\caption[TEM imaging on core-shell NWs: 4]{HRTEM imaging on one core-shell nanowire showing consumed core.}
\label{fig:ch5ws2tem2}
\end{figure}

\subsection{TEM-Raman Integrated Study on Individual NW}\label{tms:raman}

To perform TEM-Raman integrated study, the core-shell specimen was first dispersed into ethanol solution and then dipped onto polydimethylsiloxane (PDMS) support. A home-built micromanipulator with proble tip (tungsten, 100nm point radius, Micromaipulator) was used to pick up the core-shell NWs and transfer them onto TEM grids (Lacey C, 300 mesh, Cu, 01895-F Ted Pella, Inc) under 100x objective. The steps are visualized in Fig.~\ref{fig:ch5ws2trans}. 
% ws2 transfer
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{ws2trans}
\caption[Manipulating core-shell NWs]{Core-shell nanowires transfer steps. (a) tungsten tip under X100 objective, (b) NWs dispersed onto PDMS, (c) one NW picked up using tungsten tip and (d) tranferred NW on TEM grid}
\label{fig:ch5ws2trans}
\end{figure}

The transferred NWs were first examined using TEM to identify the \ce{WS2} layers configuration and the associated geometrical orientation. Then the core-shell NWs on TEM grids were probed under Raman in a location-resolved manner. Only those core-shell NWs exhibiting relative smooth \ce{WS2} wall number transition was used in subsequent Raman measurement (wall number and layer number were used interchangeably in this work, but 1 wall is not equal to 1 layer due to the tubular structure). To perform TEM and Raman analysis on individual NWs, only one NW was placed in each cell on the TEM grid (65 $\times$ 65 $\mu$m). And the low magnification TEM images and optical images were compared to map the NW location and orientation. The carbon film patterns served as a unique background to resolve possible ambiguity. An example is given in Fig.~\ref{fig:ch5ws2map}. 
% ws2 map
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{ws2map}
\caption[Mapping core-shell NWs]{Core-shell nanowires mapping.  Optical image of one NW on TEM under (a) x100 magnification and (b) digital zoom-in view; (b) Low magnification and (d) high magnification TEM images on the same NW.}
\label{fig:ch5ws2map}
\end{figure}

Fig.~\ref{fig:ch5ws2ram}a-d depict typical \ce{WO3}-\ce{WS2} NWs used in TEM-Raman integrated study. The diameter of these core-shell NWs is around 100 nm and the length around 10 $\mu$m, allowing for up to five distinctive Raman scattering sites. Fig.~\ref{fig:ch5ws2ram}e shows the associated Raman spectra acquired on core-shell NWs using 532 nm excitation. Two strong peaks were observed between 300 and 450 \si{cm^{-1}}. The resonant nature\cite{Stacy1985} of Raman shift is recognized by the broad band at about 350 \si{cm^{-1}}.  Recent theoretical\cite{Molina-Sanchez2011} and experimental\cite{Staiger2012} studies on 2D and tubular \ce{WS2} agree this band consist of first order mode $E_{2g}^1$ and second order mode $2LA(M)$, although there is still discrepancy on the exact symmetry assignment, i.e., the presence of $E_{2g}^1(M)$ mode.\cite{Berkdemir2013,Peimyoo2013}. The other peak at about 420 \si{cm^{-1}} is assigned to $A_{1g}$ mode in both bulk\cite{Sekine1980} and few layer \ce{WS2}. Tubular \ce{WS2} structure can be viewed as rolling up the planar \ce{WS2} sheet, and new Raman features arise subsequently, \emph{e.g.}, a shoulder mode emerges on the low frequency side of $A_{1g}$ mode. Theoretical investigation suggests assigning this mode as $B_{1u}$, the Davydov doublet of $A_{1g}$ mode.\cite{Ataca2012} And experimental study confirmed this $B_{1u}$ mode arise from structural disorder of \ce{WS2} layers, and the intensity ratio between $B_{1u}$ and $A_{1g}$ strongly depended on excitation wavelength.\cite{Staiger2012} This study observed $B_{1u}$ mode under 532 nm is much less prominent than that of 632 nm, in consistent with previous reports.\cite{Krause2009,Krause2009a}

% fig Raman TEM
\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{ws2_tem_ram}
\caption[TEM-Raman integrated characterization on core-shell NWs]{(a)-(d) HRTEM images on several NWs showing the \ce{WS2} wall number variation from one to about ten. (e) Raman spectra acquired from core-shell NWs supported on TEM grids, indicating a sharp contrast between the in-plane and out-of-plane vibrations. Notice that the Raman spectra primarily arise from the outer \ce{WS2} shell, and only weak Raman shift at about 800 \si{cm^{-1}} were observed at some bare core region (not shown here). (f) Layer number dependence on averaged intensity ratio between the in-plane and out-of-plane Raman bands. }
\label{fig:ch5ws2ram}
\end{figure}

Multi-peak Lorentzian fitting was also attempted on spectra shown in Fig.~\ref{fig:ch5ws2ram}e. Similar trend was found between the apparent peak ratio and the de-convolution value, as shown in Fig.~\ref{fig:ch5ws2pr}. For a qualitative purpose, the overall intensity ratio $I_{E/A}$ between in-plane E band ($E_{2g}^1+2LA$) and out-of-plane A band ($A_{1g}+B_{1u}$) is sufficient. This value is then extracted to correlate with layer number variation. Multiple Raman spectra for each layer were averaged, as shown in Fig.~\ref{fig:ch5ws2ram}f.

% fig Raman TEM
\begin{figure}[htb]
\centering
\includegraphics[width=0.4\textwidth]{ws2_peak_ratio}
\caption[Multi-peak Lorentzian fitting on core-shell NWs]{Comparison of apparent peak ratio and de-convolution value.}
\label{fig:ch5ws2pr}
\end{figure}

The intensity ratio $I_{E/A}$ decreases monotonically with the increase of \ce{WS2} wall number, a similar trend as observed on 2D few layer \ce{WS2}. The exact value of $I_{E/A}$ shows modest difference with 2D sheet, which is attributed to the suspended configuration of current core-shell NWs specimen and the presence of \ce{WO3} core. It has been demonstrated that the optical interference from dielectric environment significantly modified the Raman scattering of layered MoS2, regardless lattice vibration modes.\cite{Li2012} The author calculated the reflectivity of three different stacking scenarios to estimate the influence of \ce{WO3} and insulating substrate, as shown in. It can be concluded that under 532 nm excitation, the \ce{WO3} core shows little influence on the Raman intensity when \ce{WS2} tube is suspended in air, whereas the \ce{SiO2}/Si substrates will dramatically enhance the Raman scattering of the core-shell NWs. It could be further predicted that current Raman spectra on core-shell NWs can be used to estimate the \ce{WS2} nanotube wall numbers.

PL was examined with core-shell NWs on insulating \ce{SiO2}-Si substrates. A typical Raman spectrum from 160 to 6000 \si{cm^{-1}} is presented in Fig.~\ref{fig:ws2ramall}. This pattern was acquired on the middle spot of the core-shell NW shown as inset. Using 532 nm excitation, the multi-phonon resonant feature was clearly resolved. A broad PL peak was also observed with center at 628 nm and FWHM of 23 nm, corresponding to a photon energy of 1.974 eV with width of 72 meV.

% ws2/wo3 on sio2,1 
\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{ws2_sio2_all}
\caption[Raman spectrum of \ce{WO3}-\ce{WS2} on \ce{SiO2}-Si 1]{Raman spectrum acquired on the middle spot of core-shell NW (inset) using 532 nm. The $x$ axis is in logarithmic scale.}
\label{fig:ws2ramall}
\end{figure}

Another feature is the presence of three peaks centering at 1860, 3695, and 5500 \si{cm^{-1}}, respectively, where the latter two is probably the secondary and third overtone of the first peak. The origin of this peak family is still under research. A possible candidate points at the hydrocarbon contamination attached onto the core-shell NWs, as already been observed using TEM imaging. Interesting the intensity of this three-peak family attenuated significantly when using 441 nm excitation, as shown in Fig.~\ref{fig:ws2ramblue}. This comparison suggested there might be some enhancement mechanism from \ce{WS2} resonant Raman vibrations.

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{ws2_sio2_all_blue}
\caption[Raman spectrum of \ce{WO3}-\ce{WS2} on \ce{SiO2}-Si 2]{Raman spectrum acquired on the middle spot of core-shell NW shown in the inset of Fig.~\ref{fig:ws2ramall}.}
\label{fig:ws2ramblue}
\end{figure}

Three Raman sites were chosen to probe possible layer variation using both 532 and 441 nm, with the naming displayed in the inset of Fig.~\ref{fig:ws2ramall}. The Raman patterns are shown in Fig.~\ref{fig:ws2ram3site}a-b. The off-resonance profile in Fig.~\ref{fig:ws2ram3site}b shows only two modes, with $E_{2g}^1$ mode at about 357 \si{cm^{-1}} and $A_{1g}$ mode at $418\sim419$ \si{cm^{-1}}. In contrast, the on-resonance profile in Fig.~\ref{fig:ws2ram3site}a exhibits much richer spectra. The shoulder mode at tip2 spectrum is a manifestation of $B_{1u}$ mode from curved \ce{WS2} layer. According to previous analysis, the apparent ratio between in-plane and out-of-plane vibration indicates site tip2 has multi-wall \ce{WS2}, and the other two possibly have few-wall \ce{WS2}.

\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:green}\includegraphics[width=0.45\textwidth]{ws2_sio2_raman_green}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:blue}\includegraphics[width=0.45\textwidth]{ws2_sio2_raman_blue}}
\caption[Raman spectrum of \ce{WO3}-\ce{WS2} on \ce{SiO2}-Si 3]{Raman scattering from (a) 532 nm and (b) 441 nm excitation scanned along the core-shell NW.}
\label{fig:ws2ram3site}
\end{figure}

This predication is confirmed by the associated PL measurement as displayed in Fig.~\ref{fig:ws2plgreen}. Strong luminescent peaks were found at site mid and tip1, and an extremely weak from site tip2. The PL peaks centered at 1958 and 1934 meV for mid and tip1, respectively. The second and third order peaks also show up on all three Raman sites.

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{core-shell-sio2-pl_green}
\caption[PL spectra of \ce{WO3}-\ce{WS2} on \ce{SiO2}-Si]{PL spectrum acquired on the core-shell NW shown in the inset of Fig.~\ref{fig:ws2ramall}.}
\label{fig:ws2plgreen}
\end{figure}

\section{Summary}