### 2014-09-28 misc/mwe.tex

\documentclass[12pt,oneside]{book}
 %
   \usepackage{uncc-thesis} % some format specifications define page number in up right corner
   \input{package.tex} % additonal packages and tweaking

   \newacronym{cvd}{CVD}{chemical vapor deposition}
   \newacronym{cft}{CFT}{crystal-field theory}
   \makeglossaries

\begin{document}

  \title{\LaTeX is great}
  \author{me}
  \maketitle

    \clearpage
    \tableofcontents
    \clearpage
    \listoftables
    \clearpage
    \listoffigures

    \begin{singlespace}
    %\thispagestyle{myheadings} % before, this does not work
    \renewcommand{\glossarypreamble}{\thispagestyle{myheadings}} % after, this works.
    %\renewcommand{\acronymname}{LIST OF ABBREVIATIONS}
    \printglossary[type=\acronymtype]
    \addcontentsline{toc}{chapter}{\acronymname}
    \clearpage
    \end{singlespace}

 %  \include{./test}
     \include{./SI_thesis}
     \include{./draft}
     \include{./paper}

   %\glsaddallunused
    \appendix
    \chapter{math}
    \renewcommand{\bibname}{REFERENCES}
    \begin{singlespace}
    \printbibliography
    \end{singlespace}

\end{document}