### 2014-09-28 main.tex

\documentclass[12pt,oneside]{book}
   \usepackage{uncc-thesis} % uncc format specification seems this style should be loaded before other packages
   \input{package_iff.tex} % additonal packages and tweaking
   \input{./FrontMatter/abbre}
   \makeglossaries
\begin{document}
    \input{./FrontMatter/FrontMatter}
    \include{./ch-intro/ch-intro}
    \include{./ch-intro/ch-instr}
    \include{./ch-wox/ch-wox}
    \include{./ch-moox/moox} % no tex needed, otherwise there is error
    \include{./ch-tms/tms}
%\appendix
  % \glsaddallunused
   %\include{./BackMatter/app1}
   %\include{./BackMatter/app2}
   %\include{./BackMatter/app_training}
   % \include{./BackMatter/app-tms}
%   \include{./BackMatter/paper}

\renewcommand{\bibname}{REFERENCES}
\begin{singlespace}
  \printbibliography
\end{singlespace}

\end{document}