### 2014-10-02 FrontMatter/FrontMatter.tex

  %\title{Controlled Growth and Property Measurement of One-dimensional Oxide Nanostructures for Energy Applications}
  \title{Controlled growth and property measurement of one-dimensional oxide nanostructures for energy applications}
  \author{Tao Sheng}
  \university{The University of North Carolina at Charlotte}
  \location{Charlotte}
  \thesisyear{\the\year}
  \degree{Doctor of Philosophy}
  \department{Physics and Optical Science}

  \committeemember{Dr. Haitao Zhang} % First sets the \advisor default.
  \committeemember{Dr. Tsinghua Her}
  \committeemember{Dr. Michael Fiddy}
  \committeemember{Dr. Stuart Smith}
  \committeemember{Dr. Yu Wang} % Maximum of 10 can be specified.

%  \advisor{Dr. Haitao Zhang}

\maketitle

\input{./FrontMatter/Abstract}

\begin{dedication}
TBD.
\end{dedication}

%\pagebreak{}

\begin{ackn}
TBD.
\end{ackn}

%\pagebreak{}
%% --- Work around some bugs due to babel package---
% for more, see http://tex.stackexchange.com/questions/82993/how-to-change-the-name-of-document-elements-like-figure-contents-bibliogr?lq=1
\setcounter{tocdepth}{2}
% ----- end of work around--------

\clearpage
\tableofcontents
\clearpage
\listoftables
\clearpage
\listoffigures

\renewcommand{\glossarypreamble}{\thispagestyle{myheadings}}
\renewcommand{\acronymname}{LIST OF ABBREVIATIONS}
\glossarystyle{superborder}
%\glsaddall
\setlength{\glsdescwidth}{0.7\textwidth}
\printglossary[type=\acronymtype, toctitle=\acronymname]

%\addcontentsline{toc}{chapter}{\acronymname} this method has wrong page numbering for list of abb
% more than one page; instead, use toc option when loading glossaries package, and specify % toctitle when printing it.

\pagebreak{}