### 2014-10-09 ch-intro/ch-intro.tex
intro writing after format review

%\input{../header.tex} %  comment out when assembling

\chapter{introduction}

\section{Background and Motivation: Nanomaterials for Energy Applications}

Transition metal generally refers to any element in the d-block of periodic table, and \glspl{tmo} mean the chemical compounds that consist of at least one transition metal element and oxygen. Similarly, \glspl{tmdc} are compounds that combine at least one transition metal with chalcogen elements from group VIA (S, Se, Te, Po).\footnote{Oxygen also belongs to group VIA, but is often treated separately.} This study focused on the following three materials: molybdenum oxide, tungsten oxide and tungsten disulfide. Both tungsten (W) and molybdenum (Mo) belong to Group VIB transition metal with outer shell electrons configuration as $4d^55s^1$ and $5d^46s^2$, respectively.

The two metal oxides, \ce{MoO3} and \ce{WO3}, have been studied for decades for energy related applications, e.g., photoelectrochemistry,\cite{Su2010} photocatalysis,\cite{Watcharenwong2008, Macphee2010} photovoltaics\cite{Coridan2013} and electrochromism.\cite{Yoshimura1985, Mortimer2011} \ce{WS2} has also been investigated in various forms, i.e., thin film, nanotube, and 2D atomic layer recently. \ce{WS2} exhibits potentials in solar cell, water cleavage, and field effect transistor.  Heterostructures formed between \gls{tmo} and \gls{tmdc} provide another degree of freedom to fine tune the band structures, which could potentially enhance the photon energy harvesting applications.

\subsection{Electrochromism for Energy Saving}

Commercial buildings in U.S.\ consumed approximately 35\% of the electricity use.\footnote{U.S. Department of Energy, "Energy Efficiency Trends in Residential and Commercial Buildings", 2008.} And traditional windows in the buildings account for up to 40\% of the total heating, cooling and lighting consumption, which is of low energy efficiency. The smart window is an energy efficient answer by allowing for separate control of the visible and NIR illumination/transmission. Several solutions have been proposed in both academia and industry to develop smart window, including liquid crystal, suspended particles and electrochromic materials. Previous efforts show the electrochromic materials is the most promising component in smart window design in comparison to other options, in that it requires little power to switch between bleaching and coloration status (changing the optical transmission) and the failure mode is being clear instead of opaque, which is of great importance in commercial application.

Electrochromism, where the optical property is changed upon applying low DC voltage (\textless 5 V), have been found in numerous inorganic and organic substances, e.g., \ce{V2O5}, \ce{WO3}, and \ce{NiO} for former group, and prussian blue, polyaniline for the latter one. This study covers two inorganic metal oxides, i.e., \ce{WO3} and \ce{MoO3}. Detailed electrochromic mechanism will be discussed in Chapter 1; a brief review of electrochromism study on \ce{WO3} is presented here.

Typical transmission spectra of prototype \gls{ecd} based on \ce{WO3} thin film are shown in Fig.~\ref{}, where T in bleaching and coloration states 
A thin film example on the valency number, color center density. some calculation on doping, uv-vis results on the thin film. 
how the transmission, absorption, and reflection is modulated.

The problems of amorphous wo3.

\gls{tmo} are a group of functional materials whose bleaching-coloration status can be reversibly switched using a finite durance of low DC voltage (\textless 5 V). Another important advantage of smart windows based on EC materials is the failure mode being clear instead of opaque compared to other techniques.

exhibits electrochromic functionality. reversible, recoverable, electrochromism, 
coloration efficiency, the doping amount, 
ion mobility is the bottleneck, 
The optical properties modification in EC materials is related to the electron density difference.

\subsection{Photocatalysis for Solar Energy Harvesting}

Photocatalytic activity occurs when a semiconductor in aqueous solution is illuminated by photons of energy larger than the bandgap, then the carriers (electron and hole) drift and generate free radicals, i.e., \ce{OH.}, which enable further reactions. For water splitting using solar energy, the bandgap should be within $2.0 \sim 3.0 $ eV, and CB edge should be more negative than reduction potential of \ce{H^+/H2}, whereas the VB top should be more positive than the oxidation potential of \ce{H2O}/\ce{O2}.\cite{Osterloh2008,Wang2012} The valence band holes (\ce{h^+}) oxidize water to oxygen and conduction band electrons propel hydrogen generation, as depicted in Eq.~\ref{eq:h2o}.
\begin{subequations}\label{eq:h2o}
\begin{align}
4\ce{h^+} +  \ce{H2O} &\rightarrow \ce{O2} + 4\ce{H^+} \label{eq:hervb}\\
4\ce{H^+} +  4e^- &\rightarrow 2\ce{H2} \label{eq:hercb}
\end{align}
\end{subequations}
As shown in Fig.~\ref{fig:woxnhe}, \ce{WO3} has \gls{cb} edge positioned slightly more positive than reduction potential of \ce{H^+/H2} versus \gls{nhe}, and \gls{vb} edge much more positive than the oxidation potential of \ce{H2O}/\ce{O2}. So the photo-cleavage of water cannot be accomplished by \ce{WO3} alone. Nevertheless, a tandem cell approach by \ce{WO3} film and dye-sensitized \ce{TiO2} has been demonstrated with an efficiency of 4.5\%.\cite{Michael1999} It is worth noting that green plants also have two photosynthetic systems connected in series, one for oxidation of water into oxygen and the other for fixation of carbon dioxide.
% wo3 vs NHE
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{woxnhe.jpg}
\caption[Band positions of \ce{WO3} versus NHE]{Band positions of \ce{WO3} in contact with aqueous electrolyte at pH 1, adapted from Ref.\cite{Gratzel2001}}
\label{fig:woxnhe}
\end{figure}
Moreover, favorable oxygen evolution of \ce{WO3} brings good performance in degradation of organic compounds\cite{Hepel2001,Luo2001,Watcharenwong2008}. The formation of long-lived holes is recognized as a key requirement.\cite{Pesci2011} Besides, \ce{WO3} is remarkably stable in acid, making it a significant candidate for treating water pollutant caused by organic acids.\cite{Monllor-Satoca2006}

For photochemical water reduction to occur, the flat-band potential of the semiconductor (for highly doped semiconductors, this equals to the bottom of the conductance band) must exceed the proton reduction potential of 0.0 V vs NHE at pH=0.\cite{Osterloh2008} Flat-band potentials strongly depend on ion absorption (protonation of surface hydroxyl groups), crystallographic orientation of the exposed surface, surface defects, and surface oxidation processes.

\subsection{Advantages of Nano-Engineering}

why nano? surface-to-volume ratio, more surface area for catalytic reaction; surface energy state: tuned by dimension; quantum confinement effects: exciton size vs physical dimension. easy for dopant diffusion, thereby band structure modification; charge-separation and transport mechanism may also differ from bulk.

how to show the crystalline TMO is better than amorphous thin film counterparts.

In comparison to oxides thin films, 1D \gls{tmo} nanostructures (e.g., nanowires) have following advantages to improve the response speed and enhance the coloration efficiency: (1) With reduced diameters down to nanometer scale, the distance for ion diffusing into and out of the nanowires is significantly shortened. (2) The gap between nanowires always provides channels for fast ion diffusion outside the nanowires. (3) The large surface area to bulk volume ratio of nanowires is desired to improve the overall usage of material, which results in enhanced the contrast between colored and bleached states. Furthermore, doped with certain metals (such as Ti and Ag), the crystalline structure of the TMO nanowires can be modified with enhanced electrochromic properties.\cite{Xiong2008}

\section{Objective and Scope}
To fully harvest the advantages of these \gls{tmo} and \gls{tmdc} one-dimensional nanomaterials, a scalable growth method and a comprehensive understanding towards the growth-structure-property relation must first be established. Considering the nature of this Ph.D.\ study, this dissertation primarily focuses on synthesis, growth mechanism study and characterization, and will only present preliminary results on device and applications.

\section{Dissertation Outline}
The materials studied in this dissertation are tungsten oxides (\ce{WO3}), molybdenum oxides (\ce{MoO3}),and tungsten disulfide (\ce{WS2}). Experimental tools will be introduced in Chapter 2, including CVD apparatus and substrate treatments for growth, electron microscopes and optical spectrometers for characterization. Chapter 3 is devoted to tungsten oxides (\ce{WO3}) and sodium tungsten oxides (\ce{Na5W14O44}) nanowires, where the role of sodium in metallic tungsten precursors are thoroughly investigated; a potentially scalable growth method is also developed. Chapter 4 will turn to molybdenum oxide (\ce{MoO3}) growth using alkaline metal compounds as catalysts. In Chapter 5, a heterostructure based on \ce{WO3} \glspl{nw} is studied. Emphasis will be placed on the characterization of individual \ce{WO3}-\ce{WS2} core-shell \gls{nw}. The dissertation is then concluded with a summary and future work.