### 2015-01-10 ch-intro/ch-intro.tex

%\input{../header.tex} %  comment out when assembling

\chapter{introduction}

\section{Background and Motivation: Nanomaterials for Energy Applications}

Transition metals generally refer to any element in the d-block of periodic table, and \glspl{tmo} mean the chemical compounds that consist of at least one transition metal element and oxygen. Similarly, \glspl{tmdc} are compounds that combine at least one transition metal with chalcogen elements from group VIA (S, Se, Te, Po).\footnote{Oxygen also belongs to group VIA, but is often treated separately.} This study focused on the following three materials: molybdenum oxide, tungsten oxide, and tungsten disulfide. Both tungsten (W) and molybdenum (Mo) belong to Group VIB transition metal with outer shell electrons configuration as $4d^55s^1$ and $5d^46s^2$, respectively.

The two metal oxides, \ce{MoO3} and \ce{WO3}, have been studied for decades for energy related applications, e.g., photoelectrochemistry,\cite{Su2010} photocatalysis,\cite{Watcharenwong2008, Macphee2010} photovoltaics\cite{Coridan2013} and electrochromism.\cite{Yoshimura1985, Mortimer2011} \ce{WS2} has also been investigated in various forms, i.e., thin films, nanotubes, and 2D atomic layers recently. \ce{WS2} exhibits potentials in solar cells,\cite{Britnell2013} water cleavage,\cite{Ballif1999} and field effect transistors.\cite{Perkins2013} Heterostructures formed between \gls{tmo} and \gls{tmdc} provide another degree of freedom to fine tune the band structures, which could enhance the photon energy harvesting performances.

\subsection{Electrochromism for Energy Saving}

Commercial buildings in U.S.\ consumed approximately 35\% of the electricity usage.\cite{us2008} And traditional windows in the buildings account for up to 40\% of the total heating, cooling, and lighting consumption, which is of low energy efficiency. Smart window is an energy efficient answer by allowing for separate control of the visible and NIR illumination/transmission. Several solutions have been proposed in both academia and industry to develop smart windows, including liquid crystal, suspended particles, and electrochromic materials.\cite{Lampert1998} Previous efforts\cite{Deb2008} show the electrochromic materials is the most promising component in smart window design in comparison to other options, in that it requires little power to switch between bleaching and coloration status (changing the optical transmission) and the failure mode is being clear instead of opaque, which is of great importance in commercial applications.

Electrochromism, where the optical property of the materials is changed upon applying low DC voltage (\textless 5 V), has been found in numerous inorganic and organic substances, e.g., \ce{V2O5}, \ce{WO3}, and \ce{NiO} for the former group, and prussian blue and polyaniline for the latter one. This study covers two inorganic metal oxides, \ce{WO3} and \ce{MoO3}. Detailed electrochromic mechanism will be discussed in Chapter 1; a brief review of electrochromism study on \ce{WO3} is presented here. 
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:introecd}\includegraphics[width=0.45\textwidth]{ecd_bw}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:introtran}\includegraphics[width=0.45\textwidth]{ECD_trans}}
\caption[ECD prototype and its transmission properties]{ECD prototype and its transmission properties. (a) Schematic diagram of \gls{ecd}, and (b) optical transmission in color and bleached states, reproduced from \cite{Lampert1998}}
\label{fig:introec}
\end{figure}

Prototype \gls{ecd} usually consists of a stacked configuration as shown in Fig.~\ref{fig:introecd}. Typical transmission spectra of the prototype \gls{ecd} based on \ce{WO3} thin film are shown in Fig.~\ref{fig:introtran},\cite{Lampert1998} where optical transimission $T$ in bleaching and coloration states exhibit dynamic changes. The reflectance spectra of \ce{WO3} \gls{ecd} strongly depends on the device design.\cite{OBrien1999} As to the absorption, it was deduced that the infrared part was absorbed to a large extent, which was radiated again. Assuming the emission property of \gls{ecd} being that of blackbody, Fig.~\ref{fig:introbb} displays the wavelength distribution of spectral radiance, where the spectra span from 3 $\mu$m up to 30 $\mu$m, with peaks around $8\sim10$ $\mu$m. Further thermal engineering is needed to better regulate the infrared radiation.

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{blackbody1}
\caption[Spectral radiance of blackbody at 0, 50 and 100 \si{\degreeCelsius} predicated by Planck's Law]{Spectral radiance of blackbody at 0, 50 and 100 \si{\degreeCelsius} predicated by Planck's Law.}
\label{fig:introbb}
\end{figure}

The successful deployment of \gls{ecd} based on \ce{WO3} relies on several factors:\cite{Granqvist2000}
\begin{itemize}
\item low cost transparent conductive materials with low resistivity;
\item low cost ion electrolyte of poor electron mobility;
\item high yield growth method of \ce{WO3} and integration with other manufacturing process; 
\item life cycle $> 10^5$ for long time outdoor usage.
\end{itemize}
The deficiency of \gls{ecd} based on amorphous or polycrystalline \ce{WO3} film is the accumulation of ions at the film surface preventing further intercalation,\cite{Dini1996} which led to low coloration efficiency and short life time. In contrast, nano-structured single crystalline \ce{WO3} could bypass this limitation, as detailed in Sec.~\ref{sec:nanoadv}.

\subsection{Photocatalysis for Solar Energy Harvesting}

Photocatalytic activity occurs when a semiconductor in aqueous solution is illuminated by photons of energy larger than the bandgap, then the carriers (electron and hole) drift and generate free radicals, i.e., \ce{OH.}, which enable further reactions. For water splitting using solar energy, the bandgap should be within $2.0 \sim 3.0 $ eV, and \gls{cb} edge should be more negative than reduction potential of \ce{H^+/H2}, whereas the \gls{vb} top should be more positive than the oxidation potential of \ce{H2O}/\ce{O2}.\cite{Osterloh2008,Wang2012} The valence band holes (\ce{h^+}) oxidize water to oxygen and conduction band electrons propel hydrogen generation, as depicted in Eq.~\ref{eq:h2o}.
\begin{subequations}\label{eq:h2o}
\begin{align}
4\ce{h^+} +  \ce{H2O} &\rightarrow \ce{O2} + 4\ce{H^+} \label{eq:hervb}\\
4\ce{H^+} +  4e^- &\rightarrow 2\ce{H2} \label{eq:hercb}
\end{align}
\end{subequations}
As shown in Fig.~\ref{fig:woxnhe}, \ce{WO3} has \gls{cb} edge positioned slightly more positive than reduction potential of \ce{H^+/H2} versus \gls{nhe}, and \gls{vb} edge much more positive than the oxidation potential of \ce{H2O}/\ce{O2}. So the photo-cleavage of water cannot be accomplished by \ce{WO3} alone. Nevertheless, a tandem cell approach by \ce{WO3} film and dye-sensitized \ce{TiO2} has been demonstrated with an efficiency of 4.5\%.\cite{Michael1999} It is worth noting that green plants also have two photosynthetic systems connected in series, one for oxidation of water into oxygen and the other for fixation of carbon dioxide.
% wo3 vs NHE
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{woxnhe.jpg}
\caption[Band position of \ce{WO3} versus NHE]{Band position of \ce{WO3} in contact with aqueous electrolyte at pH 1. Reproduced from Ref.\cite{Gratzel2001}}
\label{fig:woxnhe}
\end{figure}
Moreover, favorable oxygen evolution of \ce{WO3} brings good performance in degradation of organic compounds.\cite{Hepel2001,Luo2001,Watcharenwong2008} The formation of long-lived holes is recognized as a key requirement.\cite{Pesci2011} Besides, \ce{WO3} is remarkably stable in acid, making it a significant candidate for treating water pollutant caused by organic acids.\cite{Monllor-Satoca2006}

For photochemical water reduction to occur, the flat-band potential of the semiconductor (for highly doped semiconductors, this equals to the bottom of the conductance band) must exceed the proton reduction potential of 0.0 V vs NHE at pH 0. Flat-band potentials strongly depend on ion absorption (protonation of surface hydroxyl groups), crystallographic orientation of the exposed surface, surface defects, and surface oxidation processes.\cite{Osterloh2008}

\subsection{Advantages of Nano-Engineering}\label{sec:nanoadv}

Several differences are noted in regarding to bulk materials and nanostructured materials. Nanostructured materials possesses gigantic surface area in comparison to its bulk counterpart of the same volume; for instance, given 1 gram \ce{SiO2} nanoparticles in diameter of 10 nm and assuming the density being 1.2 \si{g\per cm^3}, the estimated surface area will be $1\times10^6$ \si{cm^2}. This huge surface-to-volume ratio is generally a favorable feature for catalytic reactions, which predominately occur on the surface.\cite{Sha2009} Other features due to nanostructure include 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
\item surface energy state, thereby band structure modification;
\item quantum confinement effect when exciton size is smaller then physical dimension; and
\item different charge-separation and transport mechanism.
\end{enumerate*} Overall, these features usually promise enhanced performance.

In regarding to the contents of this study, electrochromic applications based on \gls{tmo} nanowires will benefit from the reduced diffusion depth by eliminating the deficiency of ions accumulation accounted in amorphous oxide thin film devices, which, in turn, leads to shorter response time and higher coloration efficiency.\cite{Huang2008a,Scherer2012} And photocatalytic applications using \gls{tmo} nanostructures could harvest the huge surface area on which unsaturated bond reside, thereby enabling a much higher efficiency.\cite{Mills1993,Merki2011,Chen2011b}

To make fully use of the nano-engineering, one must just be as aware of the limitations as one is of its advantages. For example, an obvious disadvantage is the difficulty of device fabrication using nanomaterials due to the size and requirements of new integration process. There also exist some uncertainty as to the energy band modification present in nanostructures, which could be either useful or not depending on the application scenario.\cite{Wang2012} With this explained, the dissertation moves on to the objective and scope section.

\section{Scope and Dissertation Outline}
To fully harvest the advantages of these \gls{tmo} and \gls{tmdc} one-dimensional nanomaterials, a scalable growth method and a comprehensive understanding towards the growth-structure-property relation must be established in the first place. This Ph.D.\ study primarily focuses on \ce{WO3} and \ce{MoO3} nanostructure synthesis, the growth mechanism study, crystal structure characterization, and optical properties measurement.

The materials studied in this dissertation are tungsten oxides (\ce{WO3}), molybdenum oxides (\ce{MoO3}),and tungsten disulfide (\ce{WS2}). Experimental tools will be introduced in Chapter 2, including CVD apparatus and substrate treatments for growth, electron microscopes and optical spectrometers for characterization. Chapter 3 is devoted to tungsten oxides (\ce{WO3}) and sodium tungsten oxides (\ce{Na5W14O44}) nanowires, where the role of sodium in metallic tungsten precursors are thoroughly investigated; a potentially scalable growth method is also developed. Chapter 4 will turn to molybdenum oxide (\ce{MoO3}) growth using alkaline metal compounds as catalysts. In Chapter 5, a heterostructure based on \ce{WO3} \glspl{nw} is studied. Emphasis will be placed on the characterization of individual \ce{WO3}-\ce{WS2} core-shell \gls{nw}. The dissertation is then concluded with a summary and future work.