### 2015-01-15 FrontMatter/Abstract.tex
all chapter fine tuned

\begin{abstract}
Transition metal oxides (TMOs) exhibit rich structures and useful properties, and could be used in applications of solar energy harvesting (\emph{e.g.}, photoelectrochemistry, photocatalysis, and photovoltaics) and energy saving (\emph{e.g.}, electrochromism and photochromism). Nano-engineering of these TMOs could produce a variety of nanostructures, modify the electronic and optical properties, and potentially enhance the performances and extend their applications into new regimes. To fully harvest the advantages of these materials, a scalable growth method and a comprehensive understanding towards the growth-structure-property relation must be established. This Ph.D. study focused on one-dimensional (1D) nanostructures of tungsten trioxide (\ce{WO3}) and molybdenum trioxide (\ce{MoO3}). This dissertation synthesized 1D \ce{WO3} and \ce{MoO3} nanostructures using chemical vapor deposition (CVD) approach, characterized the morphologies and structures, and measured the optical properties of the as-prepared nanostructures.

This dissertation work consists of three main parts, each centering on \ce{WO3}, \ce{MoO3}, and \ce{WS2}-\ce{WO3}, respectively. Systematical investigation of \ce{WO3} nanostructure using CVD method with tungsten powders as precursor led to the following accomplishments:
\begin{enumerate*}[label=\itshape\alph*\upshape)]
\item the discovery of \ce{Na5W14O44} nanowires;
\item knowledge of the delicate role of sodium in tungsten source powders; and
\item a seeded growth method to produce high quality \ce{WO3} nanowires.
\end{enumerate*} Enlightened by the intricate interactions between sodium content and tungsten oxide, this dissertation extended vapor-liquid-solid strategy into \ce{MoO3} 1D nanostructure growth, and discovered two morphologies of \ce{MoO3}: ultra-longbelts and microtowers. Careful observation in extensive experiments identified the catalytic role of sodium hydroxide applied on the growth substrate, and readily proved the catalytic behavior exists with other alkali based metal contents. The third part of this work moved the investigation center from growth to property characterization. A novel TEM-Raman integrated study was performed on \ce{WS2}-\ce{WO3} core-shell NWs, allowing for the observation of high resolution TEM imaging, Raman and Photoluminescence spectroscopy on individual nanowires. The single-walled \ce{WS2} tubular nanostructure was first identified. The combined results further illustrated the existence of \ce{WS2} wall number dependent Raman fingerprint on resonant Raman spectra.

In summary, this study explored methods for controlled and scalable growth of 1D \ce{WO3} and \ce{MoO3} nanostructures based on both vapor-solid and vapor-solid-solid approaches, characterized the structure-property relations to rationalize nanostructure synthesis for energy applications, and presented preliminary results for the devices fabrications using the \ce{WO3} and \ce{MoO3} nanostructures.
\end{abstract}