### 2015-01-15 FrontMatter/FrontMatter.tex

  %\title{Controlled Growth and Property Measurement of One-dimensional Oxide Nanostructures for Energy Applications}
  \title{Controlled growth and property measurement of one-dimensional oxide nanostructures for energy applications}
  \author{Tao Sheng}
  \university{The University of North Carolina at Charlotte}
  \location{Charlotte}
  \thesisyear{\the\year}
  \degree{Doctor of Philosophy}
  \department{Optical Science and Engineering}

  \committeemember{Dr. Haitao Zhang} % First sets the \advisor default.
  \committeemember{Dr. Tsinghua Her}
  \committeemember{Dr. Michael Fiddy}
  \committeemember{Dr. Stuart Smith}
  \committeemember{Dr. Yu Wang} % Maximum of 10 can be specified.

%  \advisor{Dr. Haitao Zhang}

\maketitle

\input{./FrontMatter/Abstract}

\begin{ackn}
\begin{singlespace}
Studying in UNC Charlotte has been an unique and rewarding experience for myself, which I shall recall frequently for a long time to come. So I am grateful to all the people here I have spent time with. Dr. Haitao Zhang has been an excellent advisor to me. I enjoyed the time learning a new set of skills, discussing research questions, and obtaining important life lessons from him. I thank Dr. Davies for the help she gave when I joined the Optical Science and Engineering program; I thank Dr. Terry Xu for allowing me to work in her lab since 2012, within which I finished most of the crystal growth work. I also thank Dr. Yong Zhang for the access to Raman instrument in his group, without which my research would be seriously diminished. I appreciate the financial supports of GASP in UNC Charlotte, the research assistantship from National Science Foundation (DMR-1006547), and the teaching assistantship from Physics and Optical Science department. I also appreciate the Center for Optoelectronics and Optical Communications at UNC Charlotte for the support in multiuser facilities. I would also like to thank all my friends and/or research partners I have met in Charlotte: Dr. Bob Cao, Dr. Jianwei Wang, Liqin Su, Henan Liu, Jinghua Ge, Siang Yee Chang, Zhe Guan, and Youfei Jiang.

Most important, I am grateful to my parents who support my decisions as always.

\begin{minipage}[b]{0.95\textwidth}

\begin{flushright}

\emph{Sheng, Tao}

\emph{January 2015}

\emph{Charlotte, NC}

\end{flushright}
\end{minipage}
\end{singlespace}
\end{ackn}

%\pagebreak{}
%% --- Work around some bugs due to babel package---
% for more, see http://tex.stackexchange.com/questions/82993/how-to-change-the-name-of-document-elements-like-figure-contents-bibliogr?lq=1
\setcounter{tocdepth}{2}
% ----- end of work around--------

\clearpage
\tableofcontents
\clearpage
\listoftables
\clearpage
\listoffigures

\renewcommand{\glossarypreamble}{\thispagestyle{myheadings}}
\renewcommand{\acronymname}{LIST OF ABBREVIATIONS}
%\glsaddall
\setlength{\glsdescwidth}{0.5\textwidth}
\glossarystyle{mysuper3col}
\printglossary[type=\acronymtype, toctitle=\acronymname]

%\addcontentsline{toc}{chapter}{\acronymname} this method has wrong page numbering for list of abb
% more than one page; instead, use toc option when loading glossaries package, and specify % toctitle when printing it.

\pagebreak{}