### 2015-02-08 FrontMatter/Abstract.tex
grammar overhaul, tense, clause, ISBN

\begin{abstract}
Transition metal oxides (TMOs) exhibit rich structures and useful properties and could be used in applications of solar energy harvesting (\emph{e.g.}, photoelectrochemistry, photocatalysis, and photovoltaics) and energy saving (\emph{e.g.}, electrochromism and photochromism). Nano-engineering of these TMOs could produce a variety of nanostructures, modify the electronic and optical properties, and potentially enhance the performances and extend their applications into new regimes. To fully harvest the advantages of these materials, a scalable growth method and a comprehensive understanding towards the growth-structure-property relation must be established. This doctoral study focused on one-dimensional (1D) nanostructures of tungsten trioxide (\ce{WO3}) and molybdenum trioxide (\ce{MoO3}). This dissertation synthesized 1D \ce{WO3} and \ce{MoO3} nanostructures using the chemical vapor deposition (CVD) approach, characterized the morphologies and structures, and measured the optical properties of the as-prepared nanostructures.

This dissertation consists of three main parts, each centering on \ce{WO3}, \ce{MoO3}, and \ce{WS2}-\ce{WO3}, respectively. Systematical investigation of the \ce{WO3} nanostructure prepared by the CVD method with tungsten powders as precursor led to the following accomplishments:
\begin{enumerate*}[label=\itshape\alph*\upshape)]
\item the discovery of \ce{Na5W14O44} nanowires;
\item a knowledge of the delicate role of sodium in tungsten source powders; and
\item a seeded growth method to produce high quality \ce{WO3} nanowires.
\end{enumerate*} Enlightened by the intricate interactions between sodium content and tungsten oxide, this dissertation extended the vapor-liquid-solid strategy into \ce{MoO3} 1D nanostructure growth and discovered two morphologies of \ce{MoO3}: ultra-longbelts and microtowers. Careful observation in extensive experiments identified the catalytic role of sodium hydroxide during the growth of \ce{MoO3} and readily proved that the catalytic behavior also exists with other alkali-based metal contents. The third part of this work moved the investigation center from growth to property characterization. A novel TEM-Raman integrated study was performed on \ce{WS2}-\ce{WO3} core-shell NWs, allowing for the observation of high resolution TEM imaging, Raman spectroscopy, and photoluminescence spectroscopy on individual nanowires. The single-walled \ce{WS2} tubular nanostructure was first identified. The combined results further illustrated the existence of \ce{WS2} wall number dependent Raman fingerprints on resonant Raman spectra.

In summary, this study explored methods for controlled and scalable growth of 1D \ce{WO3} and \ce{MoO3} nanostructures based on both vapor-solid and vapor-solid-solid approaches, characterized the structure-property relations to rationalize nanostructure synthesis for energy applications, and presented preliminary results for the device fabrications using the \ce{WO3} and \ce{MoO3} nanostructures.
\end{abstract}