### 2015-02-08 ch-moox/moox.tex

% git version control added 012614
\chapter{Vapor-Solid-Solid Growth of MOLYBDENUM OXIDE}

\section{Introduction}

In this work, a new growth mechanism of molybdenum oxide (\ce{MoO3}) 1D structures was discovered, highlighted by \ce{MoO3} long nanobelts and micro-scale towers. The samples were synthesized using a group of alkali metal based catalysts, including \ce{NaOH}, \ce{KI}, and \ce{Na2CO3}. In contrast to the sole axial growth found in the conventional catalyst-assisted growth, two different growth modes were observed for the \ce{MoO3} 1D growth here: transverse growth and axial growth. In the transverse mode, the 1D structures grew perpendicularly to the catalyst-deposition interface with catalyst particles on the side surfaces; whereas in the axial mode, the crystal grew along the catalyst-deposition axis. The growth modes were explained by a modified \gls{vss} mechanism, and the growth kinetics were explored in details.

The remaining sections are organized as following: a brief review on previous studies of \ce{MoO3} is given first, followed by the synthesis method in this dissertation. Emphasis will be placed on morphological and crystal structure characterization of 1D \ce{MoO3} nanostructures and the verification of proposed VSS mechanism. This chapter is concluded with a summary of VSS growth and preliminary results on morphology controlled \ce{MoO3} nanostructures.

\subsection{Properties and Applications of Molybdenum Oxide}

Molybdenum oxide (\ce{MoO3}) crystallizes in three phases: orthorhombic $\alpha$-\ce{MoO3}, monoclinic $\beta$-\ce{MoO3}, and the metastable hexagonal h-\ce{MoO3}.\citep{Deb1968,Fibers2007} $\alpha$-\ce{MoO3} phase (hereafter \ce{MoO3}) exhibits anisotropic structure with strong bonding along [001] and [100] direction while van der Waals interaction along [010] direction.\cite{He2003} Due to this unique structure, \ce{MoO3} has been found to have several important properties and a wide range of technological applications, such as electrochromism and photochromism,\cite{Yao1992} lubricants,\cite{Sheehan1996} photocatalysts,\cite{Chen2010} and gas sensor, including \ce{CO},\cite{Comini2005} \ce{NO2},\cite{Taurino2006} \ce{H2},\cite{Sha2009} and ethanol.\cite{Choopun} Moreover, other features arise when nanoscale \ce{MoO3} is specifically prepared, e.g. field emission.\citep{Li2002d,Zhou2003b}  As a layered hosting material, \ce{MoO3} can be further modified by intercalating with alkali ions\citep{Spahr1995,Li2006b,Hu2011} and even divalent ion.\cite{Sian2005} This structural richness has enabled improved performance in Li-ion\cite{Mai2007} and sodium-ion battery.\cite{Hariharan2013} In combination with \ce{TiO2} forming a core-shell structure, these nanoparitcles were reported to lower the photon absorption energy of \ce{TiO2}.\cite{Elder2000} When \ce{MoO3} was combined with Ag as layered structure, transparent conducting behavior was observed.\cite{Nguyen2012} In addition, \ce{MoO3} has been a good precursor for preparing other useful materials, such as \ce{MoS2} fullerene\cite{Li2003c} and few layer \ce{MoS2}.\cite{Lin2012} Recently, 2D atomic layers attracts intensive research efforts, and \ce{MoO3} can be an important member in fabricating the van der Waals heterostructures.\cite{Geim2013}

The coordination number of Mo in \ce{MoO3} is six; thus \ce{MoO6} octahedra are often considered as the building block. As shown in Fig.~\ref{fig:mo3model}, the layered structure consists of zig-zag rows of edge-sharing \ce{MoO6} octahedra, and the rows are mutually connected by corners.
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{MoO3_model}
\caption[\ce{MoO3} crystal model]{\ce{MoO3} crystal model.}
\label{fig:mo3model}
\end{figure}
Another point of view emerges when one considers the fact that four of the six surrounding O atoms are at distances from 1.67 to 1.95 \AA, while the remaining two are as far as 2.25 and 2.33 \AA, \ce{MoO3} could also be considered as chains of \ce{MoO4} tetrahedra connected by the sharing of two oxygen corners with two neighbouring tetrahedra in $c$ axis. The infinite chains of \ce{MoO4} tetrahedra form half-layers in the $ac$ plane. Two half-layers, which are stapled along $b$ axis, build up one \ce{MoO3} layer.\cite{Itoh2001a} This view stresses that the \ce{MoO6} octahedra are rather distorted.
%A brief summary of crystallography, band structure and dielectric function is listed in Appendix.

It is known that a series of molybdates form when alkali metal ions are incorporated into the \ce{MoO3} lattice, and the structures are summarized in Table~\ref{tab:naxmow}. Due to the weak cation-oxygen bonding, alkali metal cations only introduce small perturbations into the energies of \ce{Mo-O} matrix in comparison to cations of other elements. And no mixing of vibrations of the cationic sublattices with that of \ce{Mo-O} lattices is anticipated. So the structural features of \ce{Mo-O} polyhedra are dominating factors affecting the vibration frequencies and thermodynamic values of the molybdates.\cite{Fomichev1992}

\begin{table}[htb]
\centering
\caption{Crystal structures of alkali metal molybdates and tungstates}\label{tab:naxmow}
\begin{tabular}{llr}
\toprule
formula & structure  &  \\
\midrule
\ce{A2O}:\ce{MO3}\textsuperscript{\emph{a}} & isolated tetrahedral \ce{MO4} anions& \\
\ce{A2O}:\ce{2MO3} & chain-type anions of \ce{MO4} and \ce{MO6} & \\
\ce{A2O}:\ce{3MO3} & chain-type anions of \ce{MO5} and \ce{MO6} & \\
\ce{A2O}:\ce{4MO3} & chain-type anions of \ce{MO6} & \\
\bottomrule

\textsuperscript{\emph{a}} A = Li, Na, K, Rb, Cs; M = Mo, W;
\end{tabular}
\end{table}

Molybdenum bronze also exhibits intriguing features; for instance, \gls{cdw} states\footnote{In \gls{cdw} states, conductivity is non-Ohmic above a threshold electric field.} was found in blue bronze \ce{K_{0.3}MoO3}, which can be prepared by electrolytic reduction of \ce{K2MoO4} and \ce{MoO3} melt.\cite{Dumas1983} \ce{K_{0.3}MoO3} stays as monoclinic phase at room temperature with lattice parameter $a=18.249$ \AA, $b=7.561$ \AA, $c=9.856$ \AA, and $\beta=117.54^{\circ}$ and exhibits a semiconductor-to-metal transition at 180 K. Electrical transport measurements yield an highly anisotropic ratio of DC conductivities of 1:10:1000. Along this high conducting axis, metallic reflection behavior is confirmed by optical measurements. Therefore, this blue bronze $\ce{K_{0.3}MoO3}$ is also known as quasi-1D metal.\cite{Sing1999}

\subsection{Review of Synthesis Methods of Molybdenum Oxide}

Researchers have been exploring a variety of methods to synthesize different \ce{MoO3} nanostructures. Only a brief introduction will be given here, since these efforts has been well documented in several review articles.\cite{He2003} Most of these methods can be categorized into two groups: solution-based hydrothermal procedures \citep{Li2002b,Xia2006,Li2006a,Camacho-Bragado2006} and chemical vapor deposition approaches.\citep{Zeng1998,Li2002c,Li2002d,Zhou2003b,Fibers2007,Yan2009} Both methods have its own merits. Hydrothermal process usually needs lower temperature(\textless 300 \si{\degreeCelsius}), but the duration is usually long, and the sample requires several post-growth processing steps; whereas vapor deposition demands relatively high temperature (\textgreater 500 \si{\degreeCelsius}) yet a shorter time. Both methods are scalable for industrial applications. However, the hydrothermal treatment seems to allow more  nanostructures control than does the vapor deposition method. Nanobelts,\cite{Li2002b} helical nanosheets, nanoflowers, prims-like rods,\cite{Li2006a} and nanoribbons\cite{Camacho-Bragado2006} were obtained by the former one while nanoflakes,\cite{Chen2009} nanobelts,\cite{Hu2009} and nanowires\citep{Zhou2003b,Chen2011b} dominated the product morphology for the latter one. The assistance of Au catalyst in vapor deposition only altered the orientation\cite{Yan2009} or served as preferred nucleation sites\cite{Cai2011} without producing new \ce{MoO3} structures. \citeauthor{Chithambararaj2013} prepared hexagonal \ce{MoO3} nanocrystal via hydrothermal method and demonstrated the photodegradation of methylene blue (MB) under visible light.\cite{Chithambararaj2013} The efficiency dependence on catalysis/dye ratio, light intensity, and temperature was studied. h-\ce{MoO3} was mostly synthesized using solution methods, where \ce{NH4+} and \ce{OH-} were possible structure directing and stable agents. The band gap of h-\ce{MoO3} estimated from diffusion reflection spectrum is $2.8\sim3.0$ eV. Hexagonal phase of \ce{MoO3} is readily identified by the XRD pattern (strong peak at $2\theta=20^{\circ}$). It is worth noting that in spite of these numerous growth of \ce{MoO3}, reports on catalytic growth are still scarce.

\section{Experimental}\label{sec:grow}

The reactants used in this study were listed in Table~\ref{tb:mosource}. All reactants were used as received without further processing.
\begin{table}[htb]
\centering
\caption{Reactants list of \ce{MoO3} growth}\label{tb:mosource}
\begin{tabular}{lcccr}
\toprule
Material & Stock No & LOT &Purity & Vendor\\
\midrule
\ce{NaOH}     & S318-500 & 070241 & 99.8\% & Fisher Scientific \\
\ce{NaI}      & 11665 & K11W054 & 99.9\% &  Alfa Aesar \\
\ce{KI}        & 42857 & H06Z051 & 99.9\% &  Alfa Aesar \\
\ce{Na2CO3}    & 33377 & 114X012 & 99.95\% &  Alfa Aesar \\
\ce{Molybdenum}& 00932 & I07S024 & 99.9\% &  Alfa Aesar\\
\bottomrule
\end{tabular}
\end{table}

Silicon substrates was first cleaned according to the procedures stated in Sec.~\ref{ch2sub} on page~\pageref{ch2sub} and directly used in non-catalyst-assisted experiment. In catalyst-assisted growth, hydrophilic Si substrates were first drop-cast with \ce{NaOH} (or \ce{KI}, \ce{Na2CO3}) solution (54 $\mu$L 10 mM) and then air dried in a chemical fume hood before experiment. Other alkali ions-containing substrates, such as glass (Fisher Scientific, microscope slide, 12-549) and indium tin oxide (ITO) coated glass (Delta Technologies, 25\si{\ohm}), were cleaned by the same routine except the absence of plasma cleaning and subsequent aqueous solution dipping. Mica was cleaved right before growth without other treatment. 
\begin{table}[htb]
\centering
\caption{Physical constants of reactants }\label{tb:mothermo}
\begin{tabular}{lccr}
\toprule
Material & MP(\si{\degreeCelsius}) & BP(\si{\degreeCelsius}) & Reference\\
\midrule
\ce{NaOH}        & 318 & 1388 & \cite{crc1977}  \\
\ce{NaI}         & 651 & 1300 & \cite{crc1977}    \\
\ce{KI}          & 681 & 1330 & \cite{crc1977}   \\
\ce{Na2CO3}      & 851 & Not determined & \cite{crc1977}    \\
\ce{Na2MoO4}     & 687 & Not available & \cite{crc1977}   \\
\ce{MoO3}    & 795 & 1155 & \cite{crc1977}   \\
\ce{MoO2}    & 1100(decomp) & Not available & MSDS   \\
\bottomrule
\end{tabular}
\end{table}
In a typical catalytic synthesis (Fig.~\ref{fig:mooxgrowth}), about 2 g molybdenum powders were loaded into the uniform heating zone, and the sealed chamber was first pumped down to 10 mTorr. 
% cvd layout
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{CVD_and_temp_MoO3.jpg}
\caption[CVD growth setup of \ce{MoO3}]{CVD growth setup of \ce{MoO3}. The triangular labels $\blacktriangledown$ were measured points at ambient environment.}
\label{fig:mooxgrowth}
\end{figure}
Oxygen flow, varying between 0.1 sccm to 10 sccm, was then admitted from the upstream inlet. With 10 sccm UHP Ar as carrier gas, the overall pressure reached about 200 mTorr. The heating temperature was ramped up to 800 \si{\degreeCelsius} in 30 min and lasted for 15 to 120 min. At last, the heating power was turned off, and the chamber was allowed to naturally cool to room temperature. The substrate was placed in downstream location where the temperature was about 350 to 650 \si{\degreeCelsius} according to open air temperature profile. Non-catalytic growth shared the same procedure except the NaOH treatment on Si substrates.

\section{Results and Discussion on Molybdenum Oxide}\label{sec:result}
In this section, the growth of \ce{MoO3} on bare Si substrates was presented first, followed by catalytic growth, both of which were characterized by SEM, XRD, TEM and EDX. Then the catalytic growth mechanism was investigated using \ce{NaOH} treated Si substrate. The growth mechanism was probed using two series of control experiments. Then, a VSS growth model was proposed and verified. In the end, this growth model was applied onto other substrates with alkali metal ions, and various \ce{MoO3} nanostructures were produced.

\subsection{Non-catalytic Growth}\label{sec:nonsi}

For \ce{MoO3} non-catalytic growth, this study observed micro-flakes morphology, as shown in Fig.~\ref{fig:mosisem}. Most flakes exhibited rectangular shape with average thickness of one micron and standard deviation $\sigma_D$ of $0.34 \mu$m.
% sem in (31210-31510) batches
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:mosem1}\includegraphics[width=0.4\textwidth]{mosemsi_a}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:mosem2}\includegraphics[width=0.4\textwidth]{mosemsi_b}}
\caption[Representative morphologies of \ce{MoO3} on Si]{Representative morphologies of \ce{MoO3} on Si: (a) low magnification and (b) high magnification SEM images of representative depositions of \ce{MoO3} on Si for non-catalytic growth.}
\label{fig:mosisem}
\end{figure}

As shown in Fig.~\ref{fig:mooxch}, the crystal structure and phase of the as-synthesized specimen were examined by XRD and Raman. The XRD pattern (Fig.~\ref{fig:moxrd}) was readily indexed to the orthorhombic phase of \ce{MoO3} (ICDD PDF 05-0508, \emph{a}=3.9628 \AA, \emph{b}=13.855 \AA, and \emph{c}=3.6964 \AA). The space group is $D_{2h}^{16}(Pbnm)$. This crystal structure is indeed a unique example among transition metal oxides, representing a transitional stage between tetrahedra and octahedral coordination.\cite{Itoh2001a} The strongest peak index was (110), suggesting the orientation of \ce{MoO3} was not parallel to the substrate, in consistent with the morphology displayed in Fig.~\ref{fig:mosem1}. In contrast, different dominating orientation was observed in the XRD pattern from catalytic growth (Sec.~\ref{sec:nasi}).
% xrd raman
% 081113 sample
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:moxrd}\includegraphics[width=0.45\textwidth]{xrd_moo3}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:moram}\includegraphics[width=0.45\textwidth]{raman_moo3}}
\caption[Crystalline phase characterization of \ce{MoO3} on Si]{Crystalline phase characterization of \ce{MoO3} on Si: (a) XRD pattern and (b) Raman spectrum of typical \ce{MoO3} on Si, $\lambda_{ex} = 532$ nm.}
\label{fig:mooxch}
\end{figure}
The Raman spectrum of the as-synthesized specimen also closely matched \ce{MoO3} features in previous studies.\cite{Dixit1986,Silveira2012} During m-Raman measurement, the laser spot was carefully focused onto the \ce{MoO3} flakes, and several inspections on different positions were observed to ensure the reproductivity of spectra data. As shown in Fig.~\ref{fig:moram}, 14 distinct bands were well resolved. The 284 \si{cm^{-1}} peak represented the wagging mode for double bond \ce{O=Mo=O}. The 337 and 380  \si{cm^{-1}} peaks were assigned to \ce{O-Mo-O} bending and scissoring modes, respectively. The 199 \si{cm^{-1}} peak and two other weaker peaks at 218 and 247 \si{cm^{-1}} represented \ce{O=Mo=O} $B_{2g}$ twist, $A_g$ chain mode, and \ce{O=Mo=O} $B_{3g}$ twist mode, respectively. The 667 \si{cm^{-1}} peak was assigned to triply coordinated oxygen stretching model resulting from edge-shared oxygen in common to three octahedral. The 819 \si{cm^{-1}} peak arose from doubly coordinated oxygen stretching mode between two octahedral. The 996 \si{cm^{-1}} peak was from unshared oxygen stretching mode.\cite{Siciliano2009} The Rama symmetry assignment was summarized in Table~\ref{tab:moram}.
% raman assign
\begin{table}[htb]
\centering
\caption[Experimental Raman peaks assignment of \ce{MoO3} on Si]{Experimental Raman peaks assignment of \ce{MoO3} on Si.\cite{Eda1992,Siciliano2009}}\label{tab:moram}
\begin{tabular}{llcll}
\toprule
this work(\si{cm^{-1}}) & Sym.  &    & Assignment &   \\
\midrule
117      & $B_{2g}$    &           & $T_c$  & RCM\textsuperscript{\emph{a}}  \\
129      & $B_{3g}$    &           & $T_c$  & RCM  \\
158      & $A_g/B_{1g}$&           & $T_b$  & RCM  \\
199      & $B_{2g}$    & $\tau$    & \ce{O=Mo=O}  & twist  \\
218      & $A_g$       &           & $R_c$     & RCM  \\
247      & $B_{3g}$    & $\tau$    & \ce{O=Mo=O}  & twist  \\
284      & $B_{2g}$    &           & \ce{O=Mo=O}  & wag  \\
292      & $B_{3g}$     & $\delta$ & \ce{O=Mo=O}  & wag  \\
337      & $A_g,B_{1g}$ & $\delta$ & \ce{O-Mo-O} & bend  \\
380      & $B_{1g}$     & $\delta$ & \ce{O-Mo-O}  & scissor  \\
474      & $A_g$        & $\nu_{as}$ & \ce{O-Mo-O}  & stretch,bend  \\
667      & $B_{2g},B_{3g}$ & $\nu_{as}$  & \ce{O-Mo-O}  & stretch  \\
819      & $A_g$        & $\nu_{as}$  & \ce{O=Mo}  & stretch  \\
996      & $A_g$         & $\nu_{as}$  & \ce{O=Mo}  & stretch  \\
\bottomrule
\end{tabular}

\textsuperscript{\emph{a}} R chain mode;
\end{table}

\subsection{Catalyst-assisted Growth}\label{sec:naohsi}

All growth time of samples prepared in Sec.~\ref{sec:naohsi} was 2 h, the same as that in Sec.~\ref{sec:nonsi}. The only difference in growth conditions from Sec.~\ref{sec:nonsi} was the Si substrate treatment. When NaOH was applied onto Si, the deposition morphologies were changed significantly. Fig.~\ref{fig:ch4sem2by3} show a comparison of the morphology difference between the non-catalyst growth (Fig.~\ref{fig:ch4sem2by3}a) and the catalyst-assisted growth (Fig.~\ref{fig:ch4sem2by3}b-f) revealed by SEM imaging. The non-catalyst growth morphology has been discussed from Fig.~\ref{fig:mosisem} on page~\pageref{fig:mosisem}. Several different structures emerged when catalyst was applied, such as nanobelts in Fig.~\ref{fig:ch4sem2by3}d with length up to hundreds of microns and microtowers in Fig.~\ref{fig:ch4sem2by3}f with diameter around 10 $\mu$m. A detailed description of these \ce{MoO3} nanostructures has been published and can be found in Ref.\cite{Sheng2015}.

% sem
\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{MoO3_SEM_a.jpg}
\caption[Effect of alkali metal based catalysts on the morphology of \ce{MoO3} depositions]{Effect of alkali metal based catalysts on the morphology of \ce{MoO3} depositions: (a) SEM image of rectangular nanoplates grown without catalysts. (b-f) SEM images of different morphologies grown with NaOH catalysts: (b) nanobelts grown on top of dense array of nanoplates,(c) close-up view of the forked nanoplates from an area indicated by the square in Fig.~\ref{fig:ch4sem2by3}, (d) long nanobelts, (e) side-view of ultra-long microbelts, and (f) side-view of microtowers. Insets show detailed features of different \ce{MoO3} structures.}
\label{fig:ch4sem2by3}
\end{figure}

The crystal structures of the as-synthesized specimens were characterized with \gls{xrd} for the non-catalyst growth (Fig.~\ref{fig:ch4tem3by3}a) and catalyst-assisted growth (Fig.~\ref{fig:ch4tem3by3}b). The calculated lattice parameters and standard database value (ICDD PDF 05-0508) matched favorably, as listed in Table~\ref{tab:ch4xrd}. No other phases were found from XRD spectra for both specimens. This result indicated that if there were any catalyst phases present in the catalyst-assisted specimen, they must be of extremely low amount as compared to that of the \ce{MoO3} phase, presumably below the detection limit. 
% xrd 
\begin{table}[htb]
\centering
\caption{Lattice parameters of \ce{MoO3} derived from XRD}\label{tab:ch4xrd}
\begin{tabular}{lccc}
\toprule
           & $a$ (\AA) & $b$ (\AA) & $c$ (\AA)   \\
\midrule
Non-catalytic growth  & 3.963    & 13.882 & 3.701  \\
Catalytic growth  & 3.967    & 13.865 & 3.701  \\
PDF 05-0508       & 3.9628    & 13.855 & 3.6964  \\
\bottomrule
\end{tabular}
\end{table}
The XRD pattern of catalyst-assisted growth shown in Fig.~\ref{fig:ch4tem3by3}b was dominated by the (0$k$0) family peaks, showing a different preferred orientation from the non-catalytic growth. The reason will be discussed in Sec.~\ref{sec:nasi}. 
% tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{MoO3_TEM_XRD.jpg}
\caption[XRD and TEM on \ce{MoO3} growths with and without catalyst]{XRD and TEM on \ce{MoO3} growths with and without catalyst. XRD spectra of (a) a specimen grown without catalysts and (b) a specimen grown with NaOH catalysts. (c) Low-magnification TEM image and (d) HRTEM image of nanoplate grown without catalyst. Low-magnification TEM images of (e) forked nanoplate, (f) nanobelt, and (h) microtower, and (g) HRTEM image of nanobelt. Insets show SAED patterns of different \ce{MoO3} 1D structures}
\label{fig:ch4tem3by3}
\end{figure}
Fig.~\ref{fig:ch4tem3by3}c show low magnification TEM image of a rectangular nanoplate grown without catalysts and the SAED pattern. The nanoplate was exfoliated by ultra-sonication in acetone solution to obtain high-resolution TEM (HRTEM) image; hence, the nanoplate was broken into fragments. The SAED pattern in [010] axis and the HRTEM image (Fig.~\ref{fig:ch4tem3by3}d) show the nanoplate had a (010) top surface and two orthogonal edges along [100] and [001], respectively. For the catalyst-assisted growth, Fig.~\ref{fig:ch4tem3by3}e and Fig.~\ref{fig:ch4tem3by3}f-g confirmed both the forked nanoplate and the nanobelt had a (010) top surface and a growth direction of [001]; while Fig.~\ref{fig:ch4tem3by3}h revealed the microtower grew with (010) layers stacking along a [010] direction. Lattice measurements in Fig.~\ref{fig:ch4tem3by3}d and g indicated a planar distance of 0.38 nm for (100) planes and 0.36 nm for (001) planes. These values were slightly smaller than the standard data for $\alpha$-\ce{MoO3} and the results from XRD measurements. This deviation in lattice parameters may arise from the error in TEM measurements and/or the lattice distortion of \ce{MoO3} during the electron beam irradiation.\cite{Wang2004b} These results indicated that all the as-synthesized 1D structures had a layered structure. Raman measurement on the catalytic growth sample revealed similar \ce{MoO3} bands, as already displayed in Fig.~\ref{fig:mooxch} on page~\pageref{fig:mooxch}.

\subsection{Growth Mechanism Study}\label{sec:nasi}

The growth process of the 1D \ce{MoO3} nanoflakes in non-catalytic experiment can be explained by the \gls{vs} mechanism, as proposed in other vapor phase depositions of \ce{MoO3} nanostructures.\cite{Brezesinski2010, Kalantar-zadeh2010} It has been shown that \ce{MoO3} growth on the NaOH treated Si substrates exhibited dramatic differences. Previous study in our group had found that \ce{MoO3} deposition on ITO glass exhibited new morphologies that were not observed on Si substrates under the same growth conditions, yet the mechanism was not well investigated.\cite{predeep2011} A list of composition difference between glass and Si substrate show the additional elements of Na, K, and Ca, \emph{etc} in glass. Since the concentration of Na is highest among these extra elements, it is most likely to induce the new morphology growth. Therefore, a \gls{vls} mechanism may exist for the \ce{MoO3} growth when alkali metal ions are involved. To provide a proper context, \gls{vls} process is briefly introduced here.

\gls{vls} was first proposed by Wagner\cite{Wagner1964} and further developed by \citeauthor{Givargizov1975}.\cite{Givargizov1975} This method has become an important strategy for synthesizing 1D nanostructures.\cite{Lieber1998} Conventionally, a liquid eutectic droplet is formed by catalyst itself or by alloying with the growth material, acting as a trap of growth species. The growth is initialized by supersaturation of the liquid alloy, and subsequently precipitation occurs at the solid-liquid interface. The choice of catalyst usually is among the several noble metals, since they are physically active but chemically stable or inert in most growth scenarios. The growth materials span a wide range, including group IV,\cite{Hochbaum2005} group III-V,\citep{Dalacu2013, Xiao2012, Dubrovskii2011} group II-VI,\cite{Hao2006} and some metal oxides, such as \ce{ZnO},\citep{Huang2001a,Ramgir2010} \ce{MgO},\citep{HEUER1967, Nagashima2007} \ce{SiO2},\cite{Pan2002} and \ce{TiO2},\cite{Zhuge2012} to name a few. The catalysts can be Au, Pd, Pt, Ni,\cite{Xiao2012} Ti, Ga,\cite{Pan2002} and even KI. Understanding the interaction between liquid droplet and the solid interface allows for a rich engineering space to fine tune the geometry and structures of the as-grown nanostructures. For instances, the diameter of Si nanowires can be controlled by laser ablated catalyst\cite{Morales1998} or well-defined Au nanoparticles.\cite{Cui2001b} Under proper conditions, axial modulation can produce nanowire superlattices from group IV and III-V materials.\citep{Gudiksen2002,Bjork2002} Radial composition modulation has also been demonstrated by selectively suppressing VLS process, providing a robust routine for homogeneous or heterogeneous core-shell structures.\cite{Lauhon2002a}

Back to the \ce{MoO3} and alkali metal ions, a literature search found that there exists a \ce{Na2MoO4}-\ce{MoO3} phase diagram, as shown in Fig.~\ref{fig:ch4pd}.\cite{Hoermann1929} Note that the phases of the Na-Mo-O system can be written as \ce{Na2O.nMoO3}, where n = 1, 2, 3, and 4. 
%phase diagram
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{moo3_pdsmall}
\caption[Phase diagram of \ce{Na2MoO4} and \ce{MoO3}]{Phase diagram of \ce{Na2MoO4}-\ce{MoO3} reproduced from Ref.~\cite{Hoermann1929}}
\label{fig:ch4pd}
\end{figure}
Based on this phase diagram, the physical constants of reactants in Table~\ref{tb:mothermo}, and the growth conditions adopted in this study, it was predicted that the overall reactions occurred in the following sequences. First, Mo powder was oxidized, and \ce{MoO3} vapor was produced and transported to downstream by Ar carrier gas. Then, NaOH reacted with the incoming \ce{MoO3} vapor according to Formula~\ref{eq:pd}. 
\begin{subequations}\label{eq:pd}
\begin{align}
\cee{NaOH(l) + MoO3(g) &\rightarrow Na2MoO4 + H2O(g)\\
Na2MoO4 + nMoO3   &\rightarrow Na2O.(n + 1)MoO3}
\end{align}
\end{subequations}
The melting point of \ce{Na2MoO4} is 687 \si{\degreeCelsius}. Based on the open air temperature profile (Fig.~\ref{fig:mooxgrowth}), the substrate was located in temperature zone between 650 and 350 \si{\degreeCelsius}. The actual temperature distribution during growth was different from ambient measurement due to the pressure change and thermal conduction along the Si substrate. It was difficult to acquire the accurate temperatures along the substrate; however, by considering the flow rate, the actual temperature should be higher than that predicted by open air measurement.\cite{Subannajui2010} The continuous supply of \ce{MoO3} growth species pushed the phase evolution towards high \ce{MoO3} molar ratio end, with $n$ in \ce{Na2O.nMoO3} increasing from 1 to 4. Each site of \ce{Na2MoO4} continued absorb incoming \ce{MoO3}, forming a swelling particle. \ce{MoO3} precipitated from these seeds when a critical supersaturation concentration was reached, which was about 75\% in unit of mole according to Fig.~\ref{fig:ch4pd}. This ratio considerably exceeded that of the usual metal catalyst scenario, such as the Au-Si system. This high solubility accounted for several observations in current experiments: the dimensions of individual deposit was much larger than that of initial particle and the concentration of sodium in final products was extremely low. Based on this aforementioned analysis, the author proposed following hypothesis:

\emph{The reaction on NaOH-Si substrate follows the phase diagram, and \ce{Na2Mo4O13} serves as the catalyst promoting various 1D \ce{MoO3} growth.}

In a typical 2 h growth, the \ce{MoO3} weight increase was about 20 mg, which was $1.4\times 10^{-4}$ mol. This amount far exceeded the applied NaOH of $5\times 10^{-7}$ mol used in the procedure of Sec.~\ref{sec:grow} on page~\pageref{sec:grow}. To capture the early stage growth, the supply of \ce{MoO3} must be reduced in a controlled way. This can be accomplished by tuning two factors: oxygen flow rate and growth time. In this thesis, two series of controlled growth were used to probe the evolution of \ce{MoO3} deposition on NaOH-Si substrate. The first series regulated \ce{O2} flow at 0.1, 1, 3, and 10 sccm with fixed growth time of 15 min. The second series controlled the growth time at 15, 30, 60, and 90 min with fixed \ce{O2} flow of 10 sccm. Two series shared one common growth of 10 sccm \ce{O2} for 15 min.

The oxygen regulated growths were characterized using SEM, EDX, and Raman, as shown in Fig.~\ref{fig:ch4oxy}. When \ce{O2} flow was $\leq 1$ sccm, the deposition show micron size plate growth (Fig.~\ref{fig:ch4oxy}a and d). As oxygen flow became higher, the deposition amount increased dramatically, with a morphology of 1D forked plate decorated by triangular shape particles.
% sem
\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{MoO3_EDX_Raman}
\caption[Evolution of the early stage for catalyst-assisted \ce{MoO3} growth]{Evolution of the early stage for catalyst-assisted \ce{MoO3} growth: SEM images, micro-Raman spectra, and EDX spectra of the specimens grown for 15 min with different \ce{O2} flows of (a-c) 0.1 sccm, (d-f) 1 sccm, (g-i) 3 sccm, and (j-l) 10 sccm. }
\label{fig:ch4oxy}
\end{figure}
Micro-Raman scattering was used to probe the phase change during the oxygen regulated growth. The Raman spectra of sodium molybdates was featured with multiple strong peaks located between 800 and 1000 \si{cm^{-1}}, which was referred as high-frequency region hereafter. \ce{MoO3} phase show two peaks in this region with the strongest one at about 820 \si{cm^{-1}}; \ce{Na2Mo2O7} phase exhibited four peaks in this region with the strongest one moving to 937 \si{cm^{-1}}, whereas \ce{Na2Mo4O13} phase show even more peaks in the high-frequency region with a doublet at 962 and 970 \si{cm^{-1}}, respectively.\cite{Schofield2005,Saraiva2011} As predicted by the phase diagram (Fig.~\ref{fig:ch4pd}), Raman spectra identified \ce{Na2Mo2O7} phase in the 0.1 sccm growth (Fig.~\ref{fig:ch4oxy}b), and \ce{Na2Mo4O13} phase on the particles in the rest three growths (Fig.~\ref{fig:ch4oxy}e, h, and k). This observation supported the previous hypothesis about the reaction sequences on NaOH-Si substrate (Formula~\ref{eq:pd}). Element analysis performed on the Raman sites also confirmed the presence of Na, Mo, and O in particles growth and only Mo and O in the plate growth (Fig.~\ref{fig:ch4oxy}c, f, i, and l). It is worth noting that the absence of Na signal in the forked \ce{MoO3} plate indicated the doping level would be lower than the detection limit of 0.01 wt.\% if the Na doping ever occurred. It should also be pointed out the existence of \ce{Na2Mo3O10} phase is in doubt so far, and no Raman fingerprint is reported.\cite{Fomichev1992} Table~\ref{tb:moo3ram} compared the Raman spectra of sodium molybdates and \ce{MoO3} between present work and literature reports.
% moo3ram
\begin{landscape}
\begin{table}[htb]
\centering
\caption{Comparison of Raman shift (\si{cm^{-1}}) of sodium molybdates and molybdenum oxide}\label{tb:moo3ram}
\begin{tabular}{lllllllll}
\toprule
        & \ce{Na2Mo2O7}    &    &    & \ce{Na2Mo2O7}    &       &       & \ce{MoO3}   &    \\
\midrule
\cite{Saraiva2011} & \cite{Fomichev1992} & this work& \cite{Fomichev1992} & \cite{Schofield2005} & this work   & \cite{Eda1992} & \cite{Siciliano2009}   &  this work   \\
\midrule
 939    & 939/935 & 937         &        & 995    & 995         & 995    & 996       & 996   \\
 921    & 920/918 & 919         & 970    & 971    & 970         & 819    & 819       & 820   \\
 873    & 875/869 & 873         & 960    & 963    & 962         & 666    & 666       & 667   \\
 833    & 836/828 & 833         & 940    &        &             & 471    & 473       &       \\
 821    & 818/815 &             & 925    &        & 926         & 378    & 378       & 380   \\
 768    & 774/765 &             & 910    & 918    & 915         & 366    & 365       & 364   \\
 741    & 741/739 &             & 895    & 899    & 898         & 338    & 337       & 338   \\
 524    & 526/525 &             & 870    &        &             & 291    & 291       & 291   \\
 467    & 464/460 &             & 835    & 841    & 841         & 283    & 283       & 284   \\
 413    & 411/411 &             &        & 820    & 819         & 246    & 244       & 245   \\
 369    & 369/368 & 366         & 745    &        &             & 217    & 218       & 216   \\
 339    & 340/336 & 340         & 650    & 666    &             & 197    & 197       & 199   \\
 295    & 296/294 &             &        & 620    & 619         & 159    & 158       & 156   \\
 249    & 245/245 &             & 438    &        &             & 129    & 129       &       \\
 226    & 228/227 & 226         &        &        & 417         & 117    & 117       &       \\
        & 215/215 &             & 405    &        &             & 100    & 98        &       \\
 200    & 199/199 & 198         &        & 396    & 393         & 84     & 84        &       \\
 176    &         &             & 382    & 381    &             &        &           &       \\
 137    & 140/138 &             &        & 368    &             &        &           &       \\
 120    & 120/120 &             & 348    &        &             &        &           &       \\
 84     & 84/84   &             & 335    & 339    & 336         &        &           &       \\
        & 45/44   &             & 320    & 320    & 318         &        &           &       \\
 36     &         &             & 295    & 293    &             &        &           &       \\
        &         &             & 272    &        & 267         &        &           &       \\
        &         &             & 255    &        &             &        &           &       \\
        &         &             &        & 246    &             &        &           &       \\
        &         &             & 220    & 225    & 224         &        &           &       \\
        &         &             &        & 202    &             &        &           &       \\
        &         &             &        & 159    & 155         &        &           &       \\
\bottomrule
\end{tabular}
\end{table}
\end{landscape}
The morphologies of second series experiment at constant \ce{O2} flow and varied growth time were shown in Fig.~\ref{fig:ch4ev}. The 15 min and 30 min growth show that dense arrays of \ce{MoO3} forked nanoplates with a length of hundreds of microns covering the substrate areas at different temperatures. With the increase of the growth time, these nanoplates kept growing and became larger, longer, and denser. Long 1D \ce{MoO3} structures started to grow on top of the nanoplate arrays. For 60 min growth, long nanobelts appeared on top of the nanoplates. For the growth time of 90 min, the nanobelts grew longer, and some grew into microbelts. Microtowers started to grow as well.

\begin{figure}[htb]
\centering
\includegraphics[width=0.9\textwidth]{mooe_sem_ev}
\caption[Growth evolution of the second stage of catalyst-assisted growth of \ce{MoO3} 1D structures]{Growth evolution of the second stage of catalyst-assisted growth of \ce{MoO3} 1D structures: SEM images of the deposition at different growth temperatures with growth time of row 1: 15 min, row 2: 30 min, row 3: 60 min, and row 4: 90 min.}
\label{fig:ch4ev}
\end{figure}
To reveal the growth mechanisms for the long 1D structures in the second stage, specimens were carefully examined to locate the catalyst particles. Small catalyst particles ranging from several hundred nanometers to several microns were found at different locations of these long 1D structures. As shown in Fig.~\ref{fig:ch4edx1} and Fig.~\ref{fig:ch4edx2}, EDX element analysis confirmed the presence of Na, Mo, and O in these particles, indicating the particles were sodium molybdate catalysts.

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{MoO3_SFig_EDX}
\caption[SEM images and EDS spectra identifying the location, morphology, and composition of the catalyst particles on the nanobelt structures]{SEM images and EDS spectra identifying the location, morphology, and composition of the catalyst particles on the nanobelt structures: (a-b) a catalyst particle (I) in the middle of a nanobelt (II), (c-d) two catalyst particles (I and II) leading the growth of two perpendicular nanobelts.}
\label{fig:ch4edx1}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{MoO3_SFig_EDX_tower}
\caption[SEM images and EDS spectra for identifying the location, morphology, and composition of the catalysts on the microtower structures]{SEM images and EDS spectra for identifying the location, morphology, and composition of the catalysts on the microtower structures: (a-b) a tiny catalyst particle (I) on top of a small cone shaped microtower (II), and (c-d) a catalyst particle (III) on a large and short microtower (IV).}
\label{fig:ch4edx2}
\end{figure}
It could be of difficulty to observe the sodium molybdate particles on the as-synthesized belts and towers due to the following two factors, one is the consumption of catalysts, such as evaporation; the other is the unpredictable location of the catalyst. Unlike the conventional \gls{vls} growth where the catalysts were predominantly found at the tip of 1D nanostructures, the catalyst in current study were mostly found at the side wall of long nanobelts, and the location could be on the tip or several microns away from the tip. As for the microtowers, it was even rare to observe Na element. Some observations under similar controlled growth conditions was also listed as following to support the proposed growth model: 
\begin{enumerate}
\item When no Mo powder was used, NaOH still held onto substrate with some locations exhibiting etching.
\item 1 min growth show plates in high temperature end and irregular shapes in low temperature end.
\item 15 min growth show \ce{MoO3} plates parallel to substrate, with some belts emerging and growing out of the substrate plane. This was presumably due to the growth direction mismatch between different plates. Morphology will also depend on local NaOH concentration. NaOH trace mark turned into connected plates at the original locations, showing the high concentration mediated the morphology evolution. The possible catalyst solid was sitting on the top of plate layer. Some belt-like growth was visible on reaction chamber walls, suggesting the evaporation and transport of \ce{Na_xMoO3} vapor phase;
\item Using NaOH treated hydrophobic substrates, dense tower array growth was found along the edge areas or molten periphery of particles, where 5 min growth already show the initial stage of tower with diameter about 1 $\mu$ m, which increased to 10 $\mu$ m in prolonged growth.
\end{enumerate}

The growth of \ce{MoO3} on NaOH-Si can be divided into two stages delimited by the emerging of long nanobelts and microtowers. A growth model was proposed to account for the observed morphologies of \ce{MoO3} long nanobelt and tower-like structures.

\emph{
A transverse growth mode exists with growth direction perpendicular to the catalyst-deposition interface, promoting \ce{MoO3} nanoplate and nanobelt growth. A axial growth mode akin to conventional VLS process exists, and is responsible for \ce{MoO3} tower-like growth.
}

The first stage was dominated by transverse growth. The atomic steps in proposed growth model were depicted in the upper panel of Fig.~\ref{fig:ch4vls}. It has been verified that \ce{Na2Mo4O13} formed during the early stage of growth and acted as catalysts. Although the exact phase (liquid or solid) of \ce{Na2Mo4O13} was not known yet in current study, it can be deduced that the \ce{Na2Mo4O13} existed as an active site for the adsorption of incoming \ce{MoO3} growth species. When the catalyst was smaller than some critical size, there is no preferred growth direction. This changed when the precipitated solid exceeded certain critical dimension, and the morphology evolution was then determined by the relative size of already formed solid and the associated catalyst. Several pathways could occur after the accommodation of \ce{MoO3} vapor and supersaturation. \ce{MoO3} could precipitate at catalyst-deposition interface, or propagate along catalyst-environment boundary and spread on the already formed \ce{MoO3} surface. Meanwhile, the \gls{vs} process is not suppressed and \ce{MoO3} adatoms could directly condense on \ce{MoO3} surface.

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{vls_mode2}
\caption[Schematic drawing on the growth steps of the proposed growth mode]{Schematic drawing on the growth steps of the proposed growth mode.}
\label{fig:ch4vls}
\end{figure}

Some unique features of \ce{MoO3} must be taken into consideration to properly evaluate the dominate pathway(s). It has been reported that \ce{MoO3} can spontaneously spread over the surface of supports (\emph{e.g.} \ce{Al2O3}, \ce{SiO2}, and Au) to form a monolayer or submonolayer at a temperature (257 \si{\degreeCelsius}) well below the melting point.\cite{Leyrer1990} This strong spreading behavior can be explained by the solid-solid wetting process, in which the driving force was the decrease in total surface free energy.\cite{Leyrer1988} Although the detailed transport mechanisms of the spreading are still in debate, it is suggested that the high mobility of the Mo oxide species on substrates and \ce{MoO3} islands can promote the spreading and the ambient gases could further enhance the spreading.\cite{Gunther2000, Song2003} Based on above analysis and the final morphology of \ce{MoO3}, the dominance of accommodation and spreading pathway in higher temperature part of the growth sample was most possible in this study. Hereby the transverse growth mode was attributed to the strong spreading capability of \ce{MoO3}. The final morphology of observed nanoplate and nanobelt was a combined result of \gls{vs} and \gls{vls} growth. On the other hand, the tower-like structures in the second stage followed the axial growth mode in conventional VLS process. The molybdates can be evaporated from the higher temperature end and re-condense onto the lower temperature region. As illustrated in the lower panel of Fig.~\ref{fig:ch4vls}, this could initiate either belt growth or tower growth depending on the locations. The microtowers in this study primarily grew at the low temperature end of the substrate, where the spreading capability of the Mo oxide species was possibly limited. Thus the transverse growth pathway was probably suppressed and axial growth became dominant. The above analysis was substantiated by the multiple observation revealed by SEM imaging, as shown in Fig.~\ref{fig:ch4msch}.
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{moo3_sche}
\caption[SEM images for catalytic \ce{MoO3} growth]{SEM images for catalytic \ce{MoO3} growth: SEM images showing catalyst particles at different locations: (a) tip of a forked nanplate, (b) middle of a nanobelt, and (c) tip of another thin nanobelt; SEM images of microtowers with different morphologies: (d) cylindrical shape with tapered tip, (e) flat top, and (f) cone shape. }
\label{fig:ch4msch}
\end{figure}

The transport of molybdates in proposed growth model was verified by a side-by-side experiment, as schematically depicted in Fig.~\ref{fig:ch4sbs}a. Two substrates, one with NaOH catalyst and one bare substrate, were loaded side-by-side at the same location for the growth. To avoid any possible catalyst transports through surface diffusion, the two substrates were separated with several millimeters gap in between. 
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{MoO3_SFig_sidebyside}
\caption[Side-by-side growth of \ce{MoO3}]{Side-by-side growth of \ce{MoO3}. (a) Schematic drawing of the side-by-side growth with one NaOH catalyst treated substrate and one bare substrate. SEM images of the growth on the untreated substrates at different growth times: (b-e) 30 min and (f-i) 120 min. For 30 min growth: (b) low magnification SEM image showing labeled growth areas with different morphologies, (c) rectangular nanoplates grown without catalyst, (d) nanobelts grown with catalyst, and (e) forked nanoplates grown with catalyst. For 120 min growth: (f) rectangular nanoplates without catalyst, (g) forked nanoplates with catalyst, (h) microbelts with catalyst, and (i) microtowers with catalyst.}
\label{fig:ch4sbs}
\end{figure}
After the growth, besides the non-catalyst rectangular plates, catalyst induced 1D \ce{MoO3} structures were also found on the bare substrate (Fig.~\ref{fig:ch4sbs}b-i). This result clearly confirmed that the catalyst can be evaporated and transferred from the NaOH-treated substrate onto the bare one. The catalyst vapors can nucleate and form catalyst particles on the \ce{MoO3} deposition promoting the VSS growth of different 1D structures. It is worth mentioning that catalyst particles were not observed on all the 1D structures. This fact can be explained by following reasons:
\begin{enumerate*}[label=\itshape\alph*\upshape)]
\item Catalyst particles can nucleate at any locations along the 1D structures. Some particles may be invisible hiding on the backside of the 1D structures;
\item Because of the evaporation, the catalyst particles may become too small to be detected. Some particles could even disappeared terminating the catalyzed growth, which may explain the growth of the microtowers with a flat top.
\end{enumerate*} The evaporation and nucleation of catalyst not only promoted the growth of the 1D structures in the second stage growth, they also shaped the morphologies of these 1D structures. Many 1D structures in the second stage were found to have a tapered shape, such as triangular nanobelts (Fig.~\ref{fig:ch4sem2by3}b and d), microtowers with tapered tips (Fig.~\ref{fig:ch4sem2by3}f). Several factors can contribute to the tapering of the 1D structures. First, the reduction in catalyst size may induce the tapering. Due to the evaporation, the size of the catalyst particles can shrink during the growth resulting in the tapered growth. The \gls{vs} growth on the side surfaces was another possible mechanism, which promoted the radial growth of the 1D structures forming the tapered shape. Another important factor was the gradually reduced growth during the cooling process. During the cooling, the \ce{MoO3} vapor supply reduced, and the growth temperature gradually decreased. Hence, the growth rate slowed down producing the tapered tips on the top of some 1D structures.

\subsection{Growth on Alkali Metal Containing Substrates}\label{sec:glass}
In analogy to the \ce{Na2MoO4}-\ce{MoO3} binary phase diagram, other alkali metals, such as Li and K, could also form A-Mo-O alloy at elevated temperature. This fact allows for the usage of a variety of alkali metal based material as catalyst to promote \gls{vls} or \gls{vss} growth of \ce{MoO3}. Two group of experiments have been performed in this study; the first one used \ce{KI} and \ce{Na2CO3} treated Si substrate and the second one used alkali metal ions containing substrate, including ITO/glass, glass, and mica. Fig.~\ref{fig:ch4al1}a-b show the long belt growth on KI treated Si substrate and the forked plate growth on \ce{Na2CO3} treated Si, respectively. Compared to NaOH-Si growth, the belt on KI appeared with rough edge; whereas no long belt or tower was observed on the sample with \ce{Na2CO3}.

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{moo3_alt2}
\caption[Alternative catalysts for \ce{MoO3} growth]{Alternative catalysts for \ce{MoO3} growth. SEM images of \ce{MoO3} 1D structures: (a) nanobelts grown with KI on Si and (b) the forked plate growth with \ce{Na2CO3} on Si.}
\label{fig:ch4al1}
\end{figure}

Fig.~\ref{fig:ch4al2}a-c illustrated microtowers grown on ITO glass, sword-shape belts on glass, and nanoblets growth on mica, respectively. The towers on ITO glass exhibited stacking features. The sword-shape belts on glass arose from a combination of VSS and VS mechanism. The long belts found on mica (\ce{K(Al2)(Si3Al)O10(OH)2}) probably stemmed from potassium-catalyzed \gls{vls} mechanism.\cite{Hu2011}

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{moo3_alt1}
\caption[Alternative catalysts for \ce{MoO3} growth cont]{Alternative catalysts for \ce{MoO3} growth. SEM images of MoO3 1D structures: (a) microtowers grown on ITO glass, (b) triangular microbelts grown on glass, and (c) nanobelts and microbelts grown on mica.}
\label{fig:ch4al2}
\end{figure}
A rudimentary analysis on the \ce{Na+} diffusion in glass was performed to understand the growth dynamics of \ce{MoO3} on the alkali ions containing substrates. Because the alkali ions in glass can diffuse out at elevated temperature, the proposed \gls{vss} mechanism will modify the \ce{MoO3} deposition. Therefore, the mass transport of sodium ions in silica glass was first estimated to evaluate the amount of catalyst available during the growth of \ce{MoO3}. The driving force could be external field or concentration gradient, and the dynamics was governed by diffusion equation
\begin{align}\label{eq:nadiff}
\frac{\partial C}{\partial t} = D \frac{\partial^2 C}{\partial x^2},
\end{align}
where $C$ is concentration in unit of \si{mol\per cm^3}, and $D$ is diffusion coefficient in unit of \si{cm^2\per\second}. Diffusion coefficient highly depends on the overall environment in which the ions reside. Typical values of $D$ for Na ions were listed in Table~\ref{tab:mona}

\begin{table}[htb]
\centering
\caption[Mass transport of sodium ions in silica glass]{Mass transport of sodium ions in silica glass, with diffusion coefficient in unit of \si{cm^2\per\second}}\label{tab:mona}
\begin{tabular}{lcr}
\toprule
 Composition & Na diffusion coefficient  & Reference  \\
\midrule
Quartz      & $3.8\times10^{-2}\exp(\frac{-24500}{RT})$  & \cite{Rybach1967a}  \\
 \addlinespace[0.5em]
Sodalite      & $6.6\exp(\frac{-42500}{RT})$  & \cite{Sippel1963}  \\
 \addlinespace[0.5em]
Obsidian     & $4.4\times10^{-2}\exp(\frac{-22900}{RT})$  & \cite{Sippel1963}  \\
 \addlinespace[0.5em]
silicate glass & $3.1\times10^{-8}$ at 420\si{\degreeCelsius} & \cite{Jbara1995} \\
 \addlinespace[0.5em]
\ce{SiO2} glass & $1.1\times10^{-7}$ at 670\si{\degreeCelsius} &  \cite{FRISCHAT1968}\\
\bottomrule
\end{tabular}

R is molar gas constant (8.315 \si{\joule\per mol\per K}), and T is temperature in K.
\end{table}

The substrates used in this work were glass and ITO/glass; the relevant parameters were as following:
\begin{itemize}
\item dimension: $25\times10\times1$ mm
\item density: 2.567 \si{g\per cm^3}
\item Na concentration: 0.01 \si{mol\per cm^3}
\item $D$: $1.1\times10^{-7}$ \si{cm^2\per\second}
\end{itemize}
Substituting these values into Eq.~\ref{eq:nadiff}, the net flux $J$ was estimated to be
\begin{align}
J = -D \frac{\Delta C}{\Delta x} = 2.0\times 10^{-8} \si{mol cm^{-2}s^{-1}}.
\end{align}
In typical growth of 2 h, the net amount of Na diffusing out of substrate was then $1.44\times10^{-4}$ mol, about 10\% of total Na ions in the glass substrates. The diffusion length $\sqrt{Dt}$ was 190 $\mu$m. Note this result overestimated since the flux will reduce gradually. Meanwhile the molar amount of \ce{NaOH} applied on Si substrate was about $5\times10^{-7}$ mol. Considering the adjusted ratio, Na contents should be much more when using glass substrate than using NaOH-Si substrate. Another important factor was the evaporation of sodium molybdates. However, few relevant report could be found in literature. Only one was reported by \citeauthor{Kazenas2010}.\cite{Kazenas2010} The partial pressure of \ce{Na2MoO4} was calculated to be 0.2 mTorr at 800 \si{\degreeCelsius} and $3.5\times 10^{-3}$ mTorr at 670 \si{\degreeCelsius}.\footnote{$\log P(atm)= -13794/T + 6.19$}

\subsection{Optical Properties of Molybdenum Oxide}

%\input{./ch-moox/mooptics}
The optical properties of the as-grown specimen were evaluated using \gls{drs} method. Fig.~\ref{fig:modrs} show the hemisphere diffuse reflection spectrum from 800 nm to 250 nm, where the diffuse reflection $R_{dff}$ kept increasing from 27\% to 33\% when incident wavelength scanned from 800nm to about 400 nm; then, $R_{dff}$ dropped sharply to 15\% between 400 nm and 300 nm. The KMT function fitting was used to estimated the energy gap, as displayed in Fig.~\ref{fig:mokmt}. Extrapolation indicated a gap of 3.2 eV, corresponding to absorption edge at 387 nm. This matched well with the DRS data.
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:modrs}\includegraphics[width=0.45\textwidth]{moo3_DRS}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:mokmt}\includegraphics[width=0.45\textwidth]{moo3_kmt}}
\caption[DRS spectrum of \ce{MoO3} nanostructures]{DRS spectrum of \ce{MoO3} nanostructures: (a) diffuse reflection spectrum, and (b) the associated KMT function.}
\label{fig:moo3drs}
\end{figure}
Another reflection peak was found at 260 nm, illustrating there is more complex electronic band structures for \ce{MoO3} above 4 eV. As a comparison, the estimated energy gap and results from previous studies was summarized in Table~\ref{tab:mobg}.
\begin{table}[htb]
\centering
\caption{Optical band gaps of \ce{MoO3}}\label{tab:mobg}
\begin{tabular}{lccr}
\toprule
&\multicolumn{2}{c}{Band gap} \\
\cmidrule(l){2-3}
Reference & value(eV) & orientation & material states\\
\midrule
\cite{Deb1968}   & 2.96  & $E\parallel c_0$ & single crystal\\
\cite{Deb1968}   & 2.80  & $E\perp c_0$ & single crystal \\
\cite{Julien1995} & 2.8$\sim$ 3.2 & NA & films\\
 this work  & 3.2  & NA & nanoflakes\\
\bottomrule
\end{tabular}
\end{table}

\section{Summary}

In this chapter, a new growth model of \ce{MoO3} using alkali metal ions as catalysts was discovered and investigated in depth. The growth kinetics was closely inspected using a series of controlled experiments. A modified vapor-solid-solid mechanism was proposed and favorably verified. Two growth modes, i.e., transverse growth and axial growth, were highlighted, which not only improved the yield but also allowed for versatile morphology control of 1D \ce{MoO3} nanostructures. This methodology was further extended onto glass and ITO glass substrates, paving the way to a more integrated process of using \ce{MoO3} nanostructures onto various solar energy saving and harvesting applications.