### 2015-02-08 ch-tms/tms.tex

\chapter{TUNGSTEN OXIDE-TUNGSTEN DISULFIDE NANOWIRES}

In this chapter, a TEM-Raman integrated study on \ce{WO3}-\ce{WS2} core-shell nanostructures is presented. This study leverages a home-built micromanipulator apparatus to transfer the nanostructures from growth substrates to TEM grids. This transferring technique has allowed for correlating TEM and micro-Raman measurement on individual \ce{WO3}-\ce{WS2} core-shell nanowire and enhancing the understanding of structure-property relations, which would otherwise be blurred in conventional ensemble averaged methods. \ce{WO3}-\ce{WS2} heterostructure was chosen due to its potential applications in photocatalysis and hydrogen evolution.\cite{DiPaola1999,Chen2011b}

The remaining sections are arranged in the following order: a review on the growth of \ce{MS2} (M = Mo, W) \glspl{nt} and 2D \gls{fl} structures will be introduced first, followed by Raman spectroscopy summary on \ce{WS2} NTs and FL. The thesis will then focus on synthesis of \ce{WO3}-\ce{WS2} heterostructure and various characterizations on both ensemble and individual nanowires. This chapter is concluded with a brief summary and future work.

\input{./ch-tms/tmspart}

\section{Experimental}

The \ce{WO3}-\ce{WS2} core-shell nanowires were fabricated using a sulfurization process based on the \ce{WO3} NWs prepared from the seeded growth method in Chapter~3, Sec.~\ref{sec:sgfg}. For the sake of completeness, schematic layout of the two-step growth were illustrated in Fig.~\ref{fig:ch5grow}, including the seeded growth and subsequent sulfurization process.
% cvd NW growth
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{cs_growth_lbd}
\caption[Schematic diagrams of seeded growth of \ce{WO3} NW and sulfurization process]{Schematic diagrams of (a) seeded growth of \ce{WO3} NW and (b) sulfurization process. }
\label{fig:ch5grow}
\end{figure}
In a representative seeded growth (Fig.~\ref{fig:ch5grow}a), 4N5 or 5N tungsten source powders as defined in Table~\ref{tab:powder} were positioned in the upstream end of a quartz boat, and downstream about 2.5 inch the substrate was stationed. Additional tungsten seed powders (3N) were distributed evenly onto the substrate. After pumping down, 1 sccm oxygen and 10 sccm Ar were admitted into the chamber, respectively. The heating temperature was ramped up to 1000 \si{\degreeCelsius} in 30 min and lasted for 240 min. Subsequently the apparatus was allowed to naturally cool down to room temperature.

\ce{WO3}-\ce{WS2} core-shell NWs were synthesized using direct sulfurization of \ce{WO3} NWs. As depicted in Fig.~\ref{fig:ch5grow}b, the as-grown \ce{WO3} NWs were loaded into the center of heating furnace, and $\sim200$ mg sulfur (Alfa Aesar 10785, 99.5\%) was positioned just outside the upstream edge of furnace, where the maximum temperature was about 240 \si{\degreeCelsius}. After pumping down, the reaction chamber was flushed two times to expel residual air. A cold trap filled with liquid nitrogen in downstream was used to collect possible sulfur precipitation. Then, the furnace was heated to 750 \si{\degreeCelsius} in 30 min, held for 15 min, and allowed to naturally cool down to room temperature. During entire growth process, 30 sccm Ar was used as carrier gas.

\section{Results and Discussion}
\subsection{Ensemble Measurements on Core-Shell NWs Array}

The morphology of \ce{WO3} NWs in micron scale almost stayed the same after sulfurization, as shown in the inset SEM imaging of Fig.~\ref{fig:ch5ws2sem}. The length of individual nanowire were up to 15 $\mu$m, and the diameter varied from about 40 to 200 nm. The surface of the NWs were rather smooth without other attachments. The NWs array grew slightly larger than the W seeds ($\sim$12 $\mu$m), mostly due to the elongation of NW.
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{ws2_sem_edx_labeled}
\caption[Morphology and composition analysis on the as-synthesized \ce{WO3}-\ce{WS2} core-shell structures]{Morphology and composition analysis on the as-synthesized \ce{WO3}-\ce{WS2} core-shell structures. (a) Low magnification and (b) high magnification SEM images showing the NWs morphologies after sulfurization. (c) EDX spectroscopy revealing the presence of sulfur element after sulfurization.}
\label{fig:ch5ws2sem}
\end{figure}
The EDX spectrum in Fig.~\ref{fig:ch5ws2sem} revealed the presence of sulfur element, indicating the effectiveness of sulfurization. Compared to the XRD pattern on seeded growth \ce{WO3} sample, post-sulfurization pattern (Fig.~\ref{fig:ch5ws2xrd}) show a much reduced \ce{WO3} intensity with W phase becoming relatively prominent. The peaks at 23.44$^{\circ}$ and 24.36$^{\circ}$ were indexed to \ce{WO3} (020) and (200) reflection, respectively. This result as well as the SEM imaging indicated \ce{WO3} NWs preserved some long range order after the sulfurization process of 15 min at 750 \si{\degreeCelsius}. 
%fig xrd ws2 
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{ws2_cs_xrd}
\caption[X-ray diffraction spectra before and after sulfurization]{X-ray diffraction spectra before and after sulfurization. The reference spectra for $c$-W, 2H-\ce{WS2} and $m$-\ce{WO3} are included with major peaks labeled.}
\label{fig:ch5ws2xrd}
\end{figure}
This observation can also be used to account for the extreme small peak at 13.87$^{\circ}$ in the post-sulfurization spectrum. In comparison to the reference 2H-\ce{WS2} (002) peak (14.37$^{\circ}$, ICDD PDF 04-003-4478), the redshift of specimen peak here indicated an increase of lattice spacing along $c$ axis (from 6.16 \si{\angstrom} to 6.38 \si{\angstrom}). The FWHM of this (002) peak was about 1.1$^{\circ}$, giving an estimation on the crystalline size of 7.5 nm according to Scherrer equation.

Due to this small spacing, TEM was primarily utilized to observe the detailed morphology of \ce{WO3}-\ce{WS2} core-shell NWs. Fig.~\ref{fig:ch5ws2tem1}c and d show HRTEM imaging on typical tip and body region of the core-shell NW, respectively. A comparison with the pre-sulfurization specimen (Fig.~\ref{fig:woseedtem2} on page~\pageref{fig:woseedtem2}) readily revealed the formation of core-shell morphology after sulfurization. The measured layer spacing of 6.5 \si{\angstrom} appeared slightly larger than bulk 2H-\ce{WS2} value of 6.16 \si{\angstrom}, but matched well with the XRD result in Fig.~\ref{fig:ch5ws2xrd}. The sharp contrast in core region also proved the lattice structure of \ce{WO3} was not seriously distorted upon sulfurization.
% tem core shell layer
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{ws2_tem_labeled}
\caption[HRTEM imaging before and after sulfurization]{HRTEM imaging before and after sulfurization illustrating the presence of core-shell structure. (a) Low magnification and (b) HRTEM images on \ce{WO3} NW specimen from seeded growth;  HRTEM images on the core-shell nanowire of (c) tip region and (d) body region showing layer number variation. The contrast between (b) and (d) readily identify the formation of core-shell structure. }
\label{fig:ch5ws2tem1}
\end{figure}

After examining dozens of core-shell NWs, this study found the \ce{WS2} layer number at NW tip was usually the largest, reaching more than ten, and the layer number decreased to one along the NW body; but there also existed some layer number fluctuation and shape modification. Fig.~\ref{fig:ch5ws2tem3} show a smooth transition of \ce{WS2} layer number on the tip region of one NW. The layer number was about 7 on the tip corner, gradually decreased from 4 to 3 along the NW body, and stopped at 2 layers. Closer observation indicated the layer spacing was not quite uniform in the tip area, probably due to the complex curvature and bending. \ce{WS2} fullerene also show similar structures. Since the closing of the tip region could be of great complexity and the induced strain could modify Raman signal considerably, the TEM-Raman integrated measurement in this study focused on the body region of the core-shell NWs with smooth transition layer number.
% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{cs_tem_g1}
\caption[HRTEM imaging on the tip area of one core-shell nanowire]{HRTEM imaging on the tip area of one core-shell nanowire.}
\label{fig:ch5ws2tem3}
\end{figure}
As illustrated in Fig.~\ref{fig:ch5ws2tem4}, complicated \ce{WS2} encapsulation was observed on one tapered core-shell NW. This tapering probably arose from the sulfurization process but could be also due to the parent \ce{WO3} NW. The diameter varied from about 200 nm to 15 nm, and the \ce{WS2} layer number also changed in a nonlinear manner. On some turning region, the \ce{WS2} seemed penetrating into the body of primary \ce{WO3} NW. 
% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{cs_tem_g3_lbd}
\caption[HRTEM imaging on one tapered core-shell nanowire showing fluctuation of layer numbers]{HRTEM imaging on one tapered core-shell nanowire showing fluctuation of layer numbers.}
\label{fig:ch5ws2tem4}
\end{figure}

A systematic study on the sulfurization must be performed to understand the growth dynamics. The growth mechanism of \ce{WS2} nanotubes from oxides has been thoroughly studied,\cite{Feldman1998} and it was generally agreed that the sulfurization of \ce{WO3} was a diffusion limited reaction.\cite{Feldman1996} It was also suggested that the core \ce{WO3} NWs became non-stoichiometric phases upon sulfurization.\cite{Feldman1996,ZAK2009} This study recognized this possibility but also considered two factors different from previous studies: one was the usage of sulfur as precursor instead of \ce{H2} and/or \ce{H2S} as in previous reports, which accelerated the reduction of \ce{WO3} significantly due to the formation of \ce{H2O}; the other was the short reaction time and light degree of sulfurization, which probably would not distort the \ce{WO3} lattice significantly, as evidenced by the XRD and HRTEM analysis in this section. The sulfurization dynamic was then proposed as following in this thesis: the tips of \ce{WO3} NWs exhibited greater crystalline disorder, and \ce{WS2} layers on tips were folded at least along two axes, resulting a higher density of defects and thus a higher sulfurization rate; while the folding along NW body was quasi-one dimensional, subsequently the sulfurization rate could be lower. Besides, sulfurization could also be initiated from the NW body, presumably due to the presence of surface defects. As a result, this led to the uneven \ce{WS2} layer number distribution along the core-shell NWs.

% coreshell tem
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{core-shell_void_growth}
\caption[HRTEM imaging on one core-shell nanowire showing consumed core]{HRTEM imaging on one core-shell nanowire showing consumed core.}
\label{fig:ch5ws2tem2}
\end{figure}
No obviously consumed core region was observed among most of the core-shell NWs examined under TEM. An exception was shown in Fig.~\ref{fig:ch5ws2tem2}. Note the rectangular area show abrupt termination of one \ce{WS2} layer, which was tentatively attributed to stacking fault in this study.

\subsection{TEM-Raman Integrated Study on Individual NW}\label{tms:raman}

To perform TEM-Raman integrated study, the core-shell specimen was first dispersed into ethanol solution; and then the solution was dipped onto polydimethylsiloxane (PDMS) support. A home-built micromanipulator with probe tip (tungsten, 100nm point radius, Micromaipulator) was used to pick up the core-shell NWs and transfer them onto TEM grids (Lacey C, 300 mesh, Cu, 01895-F Ted Pella, Inc) under 100X objective. The steps were visualized in Fig.~\ref{fig:ch5ws2trans}. 
% ws2 transfer
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{cs_NW_transfer}
\caption[Core-shell nanowires transfer steps]{Core-shell nanowires transfer steps. (a) tungsten tip under X100 objective, (b) NWs dispersed onto PDMS, (c) one NW picked up using tungsten tip and (d) tranferred NW on TEM grid with two tips indicated by the arrows.}
\label{fig:ch5ws2trans}
\end{figure}
The transferred NWs were first examined using TEM to identify the \ce{WS2} layers configuration and the associated geometrical orientation. Then, the core-shell NWs on TEM grids were probed under Raman in a location-resolved manner. Only those core-shell NWs exhibiting relative smooth \ce{WS2} wall number transition were used in subsequent Raman measurement (wall number and layer number were used interchangeably in this work, but note that 1 wall is not equal to 1 layer due to the tubular structure). To perform TEM and Raman analysis on individual NWs, this study placed one NW only in each cell of the TEM grid (65 $\times$ 65 $\mu$m). And the low magnification TEM images and optical images were compared to map the NW location and orientation. The carbon film patterns served as a unique background to resolve possible ambiguity. An example was given in Fig.~\ref{fig:ch5ws2map}. 
% ws2 map
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{cs_nw_map}
\caption[Core-shell nanowires mapping]{Core-shell nanowires mapping. Optical image of one NW on the lacey  carbon TEM grid under (a) x100 magnification and (c) digital zoom-in view with two end indicated by the arrows, TEM images of (b) low magnification and (d) high magnification acquired on the same NW specimen.}
\label{fig:ch5ws2map}
\end{figure}

Fig.~\ref{fig:ch5ws2ram}a-d depicted typical morphologies of the \ce{WO3}-\ce{WS2} NWs in the TEM-Raman integrated study. The diameter of these core-shell NWs was about 100 nm, and the length about 10 $\mu$m, allowing for up to five distinctive Raman scattering sites. Fig.~\ref{fig:ch5ws2ram}e show the associated Raman spectra acquired on the core-shell NWs using 532 nm excitation wavelength. Two strong peaks were observed between 300 and 450 \si{cm^{-1}}. The resonant nature\cite{Stacy1985} of Raman shift was recognized by the broad band at about 350 \si{cm^{-1}}.  Recent theoretical\cite{Molina-Sanchez2011} and experimental\cite{Staiger2012} studies on 2D and tubular \ce{WS2} agreed that this band consisted of first order mode $E_{2g}^1$ and second order mode $2LA(M)$, although there was still discrepancy on the exact symmetry assignment, i.e., the presence of $E_{2g}^1(M)$ mode.\cite{Berkdemir2013,Peimyoo2013}. The other peak at about 420 \si{cm^{-1}} was assigned to $A_{1g}$ mode in both bulk\cite{Sekine1980} and few layer \ce{WS2}. Tubular \ce{WS2} structure can be viewed as rolling up the planar \ce{WS2} sheet, and new Raman features arose subsequently, e.g., a shoulder mode emerged on the low frequency side of $A_{1g}$ mode. Theoretical investigation suggested assigning this mode as $B_{1u}$, the Davydov doublet of $A_{1g}$ mode.\cite{Ataca2012} And experimental study confirmed this $B_{1u}$ mode arose from structural disorder of \ce{WS2} layers, and the intensity ratio between $B_{1u}$ and $A_{1g}$ strongly depended on excitation wavelength.\cite{Staiger2012} This study observed $B_{1u}$ mode under 532 nm was much less prominent than that of 632 nm, in consistent with previous reports.\cite{Krause2009,Krause2009a} 
% fig Raman TEM
\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{cs_temram}
\caption[TEM-Raman integrated characterization on core-shell NWs]{TEM-Raman integrated characterization on core-shell NWs. (a)-(d) HRTEM images on four NWs showing the \ce{WS2} wall number variation from one to about ten. (e) Raman spectra acquired from the same core-shell NWs supported on TEM grids, indicating a sharp contrast between the in-plane and out-of-plane vibrations. Notice that the Raman spectra primarily arise from the outer \ce{WS2} shell, and only weak Raman shift at about 800 \si{cm^{-1}} were observed at some bare core region (not shown here).}
\label{fig:ch5ws2ram}
\end{figure}

Rigid analysis of the Raman spectra must be performed to reach solid conclusion. Hence, multi-peak Lorentzian fitting was employed on the spectra in Fig.~\ref{fig:ch5ws2ram}e, and the result for 1 layer \ce{WS2} core-shell NW was displayed in Fig.~\ref{fig:ch5ws2ml}. The de-convolution peaks compared favorably with original Raman spectrum, and similar operation were used for other three spectra. Fig.~\ref{fig:ch5ws2pr} show the layer number dependence on three group of intensity ratios extracted from apparent value, de-convoluted height, and integrated intensity, respectively. Similar trend was found among all three set of data, though the integrated intensity appeared more sensitive towards layer No variation. For a qualitative purpose, the apparent intensity ratio $I_{E/A}$ between in-plane E band ($E_{2g}^1+2LA$) and out-of-plane A band ($A_{1g}+B_{1u}$) was used as a quick criterion to distinguish \ce{WS2} few layer in this study. 
% fig Raman TEM
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:ch5ws2ml}\includegraphics[width=0.45\textwidth]{ws2_1L_fitting}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:ch5ws2pr}\includegraphics[width=0.45\textwidth]{ws2_peak_ratio}}
\caption[Multi-peak Lorentzian fitting on core-shell NWs]{(a) Multi-peak Lorentzian fitting on core-shell NWs, and (b) Layer number dependence on averaged intensity ratio between the in-plane and out-of-plane Raman bands.}
\label{fig:ch5ws2prl}
\end{figure}
$I_{E/A}$ was then extracted to correlate with layer number variation on tens of core-shell NWs. As shown in Fig.~\ref{fig:ch5ws2ramap}, the mapping of \ce{WS2} layer dependent $I_{E/A}$ almost followed the same trend in Fig.~\ref{fig:ch5ws2pr} with some variation attributed to the core-shell NW diameter difference.
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{ws2_ratio_map}
\caption[Mapping on the \ce{WS2} layer dependent intensity ratio]{Mapping on the \ce{WS2} layer dependent intensity ratio. The Raman spectra were acquired on individual NW suspended on 300 mesh lacey C TEM grid.}
\label{fig:ch5ws2ramap}
\end{figure}
The intensity ratio $I_{E/A}$ decreased monotonically with the increase of \ce{WS2} wall number, a similar trend as observed on 2D few layer \ce{WS2},\cite{Berkdemir2013} and the exact value of $I_{E/A}$ show modest difference with 2D sheet, which could be attributed to the suspended configuration of current core-shell NWs specimen and the presence of \ce{WO3} core. Furthermore, these results indicated that $I_{E/A} > 2$ for single-walled \ce{WS2} core-shell NW, and when the \ce{WS2} wall number was about 3, $I_{E/A}=1$. These two values served as a quick determination of \ce{WS2} wall number when the NW was suspended.

Before using this observation in more general scenarios, one should evaluate the influence of \ce{WO3} and other dielectric substrates. It has been demonstrated that the optical interference from dielectric environment significantly modified the intensity of Raman scattering from layered \ce{MoS2}, regardless lattice vibration modes.\cite{Li2012} In this work, the author calculated the reflectivity of three different stacking scenarios to estimate the influence of \ce{WO3} and insulating substrate, as shown in Fig.~\ref{fig:ch5ws2stk}. It was concluded that under 532 nm excitation, the \ce{WO3} core show little influence on the Raman intensity when \ce{WS2} nanotube was suspended in air; whereas the \ce{SiO2}/Si substrates dramatically enhanced the Raman scattering of the core-shell NWs. It was further predicted that current Raman spectra on the core-shell NWs can be used to estimate \ce{WS2} nanotube wall numbers.

\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:stk1}\includegraphics[width=0.30\textwidth]{stacking1}}\hspace{0.03\textwidth}
\subfloat[]{\label{fig:stk2}\includegraphics[width=0.30\textwidth]{stacking2}}\hspace{0.03\textwidth}
\subfloat[]{\label{fig:stk3}\includegraphics[width=0.30\textwidth]{stacking3}}
\caption[Optical reflectance simulation on three different \ce{WS2} layer stacking configurations]{Optical reflectance simulation on three different \ce{WS2} layer stacking configurations.}
\label{fig:ch5ws2stk}
\end{figure}

\Gls{pl} was examined with core-shell NWs on insulating \ce{SiO2}-Si substrates. A typical Raman spectrum from 160 to 6000 \si{cm^{-1}} was presented in Fig.~\ref{fig:ws2ramall}. This pattern was acquired on the middle spot of the core-shell NW shown as inset. Using 532 nm excitation, the multi-phonon resonant feature was clearly resolved. A broad PL peak was also observed with center at 628 nm and FWHM of 23 nm, corresponding to a photon energy of 1.974 eV with width of 72 meV.  
% ws2/wo3 on sio2,1 
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{ws2_sio2_all}
\caption[Raman spectrum acquired on the middle spot of core-shell NW (inset) using 532 nm. The $x$ axis is in logarithmic scale]{Raman spectrum acquired on the middle spot of core-shell NW (inset) using 532 nm. The $x$ axis is in logarithmic scale.}
\label{fig:ws2ramall}
\end{figure}
Another feature was the presence of three peaks centering at 1860, 3695, and 5500 \si{cm^{-1}}, respectively, where the latter two were probably the secondary and third overtone of the first peak. It was confirmed that this three peaks belonged to Raman shift instead of PL, since similar peaks were observed using different excitation wavelength of 441 nm and 532 nm. The origin of this peak family was still under research, and one possible candidate was the \ce{C=O} bonding, which presumably came from the carbon impurity in tungsten powders as seeds. It was also possible that the \ce{C=O} bonding arose from the hydrocarbon contamination attached onto the core-shell NWs. Interestingly, the intensity of this three-peak family attenuated significantly when using 441 nm excitation, as shown in Fig.~\ref{fig:ws2ramblue}. This comparison suggested there might be some enhancement mechanism from \ce{WS2} resonant Raman vibrations. 
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{ws2_sio2_all_blue}
\caption[Raman spectrum acquired on the middle spot of core-shell NW shown in the inset of Fig.~\ref{fig:ws2ramall}]{Raman spectrum acquired on the middle spot of core-shell NW shown in the inset of Fig.~\ref{fig:ws2ramall}.}
\label{fig:ws2ramblue}
\end{figure}
Three Raman sites were chosen to probe possible layer variation using both 532 and 441 nm excitation with the naming displayed in the inset of Fig.~\ref{fig:ws2ramall}. The Raman patterns were shown in Fig.~\ref{fig:ws2ram3site}a-b. The off-resonance profile in Fig.~\ref{fig:ws2ram3site}b show only two modes: $E_{2g}^1$ mode at about 357 \si{cm^{-1}} and $A_{1g}$ mode at $418\sim419$ \si{cm^{-1}}. In contrast, the on-resonance profile in Fig.~\ref{fig:ws2ram3site}a exhibited much richer spectra. The shoulder mode at tip2 spectrum was a manifestation of $B_{1u}$ mode from curved \ce{WS2} layer. According to previous analysis, the apparent ratio between in-plane and out-of-plane vibration indicated site tip2 had multi-wall \ce{WS2}, and the other two possibly had few-walled \ce{WS2}. 
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:green}\includegraphics[width=0.45\textwidth]{ws2_sio2_raman_green}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:blue}\includegraphics[width=0.45\textwidth]{ws2_sio2_raman_blue}}
\caption[Raman scattering from (a) 532 nm and (b) 441 nm excitation scanned along the core-shell NW]{Raman scattering from (a) 532 nm and (b) 441 nm excitation scanned along the core-shell NW.}
\label{fig:ws2ram3site}
\end{figure}
This predication was confirmed by the associated PL measurement displayed in Fig.~\ref{fig:ws2plgreen}. Strong luminescent peaks were found at site mid and tip1, and an extremely weak peak were found from site tip2. The PL peaks centered at 1.958 and 1.934 eV for mid and tip1, respectively. The second and third order peaks also show up on all three Raman sites. 
\begin{figure}[htb]
\centering
\includegraphics[width=0.6\textwidth]{core-shell-sio2-pl_green}
\caption[PL spectra of \ce{WO3}-\ce{WS2} on \ce{SiO2}-Si]{PL spectrum acquired on the core-shell NW shown in the inset of Fig.~\ref{fig:ws2ramall}.}
\label{fig:ws2plgreen}
\end{figure}

\section{Summary}

A comprehensive understanding of the structure-property relation, which is essential to successful application of nanomaterials, is often hindered by the ensemble averaged characterizations in conventional methods. The present work in this chapter, to some extent, circumvented this limitation by leveraging a home-built micromanipulator apparatus to transfer the nanostructures between TEM grids and other desired substrates. Considerable insight on the \ce{WO3}-\ce{WS2} core-shell nanostructures was obtained: few-layer, even single-layer tubular \ce{WS2} nanostructure was observed with \ce{WO3} core; layer dependent Raman spectra were studied at on-resonance and off-resonance conditions; the fingerprint of few-layer \ce{WS2} was revealed; and photoluminescence from core-shell NWs was identified. These results held great values to the application of \ce{WO3}-\ce{WS2} heterostrucutres in photocatalysis and photoelectrochemical applications. More important, this technique of TEM-Raman integrated study could be readily applied to a variety of other nanomaterials, therefore opening a broad spectrum for the understanding of the structure-property relation, propelling the potential nano-engineering applications.

\chapter{SUMMARY AND FUTURE WORK}

In summary, synthesis of \ce{WO3} and \ce{MoO3} nanostructures based on CVD method were thoroughly investigated in this study. Sodium impurity effect in tungsten powder source was revealed, and the formation of \ce{Na5W14O44} nanowire was first reported. Partially enlightened by this discovery, a new VSS growth mechanism of \ce{MoO3} was found, and the growth dynamics was systematically probed. Comprehensive electron microscopy characterizations were performed on both \ce{WO3} and \ce{MoO3}. And the insight of growth kinetics can be used to establish a systematic connection between the interaction of transition metal oxide and alkali metal ions. As an example, preliminary results of \ce{WO3} NWs growth using \ce{NaOH} treated Si substrate and controlled \ce{MoO3} 1D nanostructures growth on glass were presented in Sec.~\ref{sec:nawo} and Sec.~\ref{sec:mogls}, respectively.

Besides, seeded growth of \ce{WO3} was demonstrated to be an effective and scalable approach for obtaining single crystalline 1D \ce{WO3} NWs. Based on these \ce{WO3} nanowires, \ce{WO3}-\ce{WS2} heterostructure was synthesized using a simple sulfurization process, and single-walled \ce{WS2} tubular core-shell nanowire was observed. More important, a TEM-Raman integrated approach was tested aiming at providing a location-resolved map between crystal structure and optical properties. As a result, layer number dependence of Raman spectra on \ce{WS2} tubular nanostructure was fingerprinted, and direct band gap \gls{pl} from \ce{WS2} tubular structure was observed. These results indicated a broad space for band structure tuning of \ce{WO3}-\ce{WS2}. Moreover, the integration of TEM-Raman served as an excellent tool to correlate the structure-property relationship for other 1D nanomaterials.

\section{Preliminary results of NaOH WO$_3$ growth}\label{sec:nawo}
The growth conditions almost remained the same as that of the non-seeded \ce{WO3} growth except that the substrate were located at 7.5 inch (refer to page~\pageref{sec:woxnonseed} for details). As shown in Fig.~\ref{fig:naohwsem}, dense NWs were observed on \ce{NaOH} treated substrate by SEM imaging, where the length ranged from 3 to 10 $\mu$m and diameter from 50 to 200 nm. Compared with the \ce{WO3} NWs in seeded growth, the deposition here also exhibited some plausible plate morphology.
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{naohwo3_sem.jpg}
\caption[SEM characterization of \ce{NaOH} \ce{WO3} growth]{SEM characterization of \ce{NaOH} \ce{WO3} growth. (a) Low-magnification SEM graphs showing dense \ce{WO3} NWs. (b) A high magnification view showing thin NW growth.}
\label{fig:naohwsem}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{naoh-wo3_edx}
\caption[Composition analysis on \ce{NaOH} \ce{WO3} growth NWs identifying small fraction Na content]{Composition analysis on \ce{NaOH} \ce{WO3} growth NWs identifying small fraction Na content.}
\label{fig:naohwedx}
\end{figure}
\gls{edx} analysis on the \ce{NaOH} \ce{WO3} sample was shown in Fig.~\ref{fig:naohwedx}. Tungsten, oxygen, and sodium elements were detected on the plate growth region with Na/W atomic ratio close to 0.05/1. 
%082713 sample
\begin{figure}[htb]
\centering
\subfloat[]{\label{fig:nawxrd}\includegraphics[width=0.45\textwidth]{naoh-wo3_xrd}}\hspace{0.04\textwidth}
\subfloat[]{\label{fig:nawram}\includegraphics[width=0.45\textwidth]{naoh-wo3_raman}}
\caption[Characterization of \ce{NaOH} \ce{WO3} NWs: XRD and Raman]{Characterization of \ce{NaOH} \ce{WO3} NWs: XRD and Raman. (a) XRD pattern of the as-prepared sample, and (b) Raman spectrum on NWs region.}
\label{fig:naohwxrd}
\end{figure}
Fig.~\ref{fig:nawxrd} displayed the XRD pattern on one representative sample. The phase identification turned out to be ambiguous. One candidate was orthorhombic \ce{WO3} phase (ICDD PDF 00-020-1324, \emph{a}=7.384 \si{\angstrom}, \emph{b}=7.512 \si{\angstrom}, \emph{c}=3.846 \si{\angstrom}); another one was monoclinic \ce{W10O29} phase (ICDD PDF 04-007-0501, \emph{a}=12.1 \si{\angstrom}, \emph{b}=3.78 \si{\angstrom}, \emph{c}=23.4 \si{\angstrom}, $\beta$=90.5$^\circ$). Micro-Raman scattering spectroscopy was performed on the as-synthesized sample as well. As shown in Fig.~\ref{fig:nawram}, five distinct bands were well resolved with peaks located at 230, 380, 428, 650 and 816 \si{cm^{-1}}, respectively. Another small hump was also observed at 946 \si{cm^{-1}}. This Raman spectrum deviated from typical features of \ce{WO3},\cite{Salje1975a,Dixit1986} but matched well with \ce{WO3.H2O} in one previous study;\cite{Daniel1987} however, the XRD pattern did not reveal tungsten oxide hydrates feature.\footnote{ICDD PDF 04-011-1708, \emph{a}=7.35 \si{\angstrom}, \emph{b}=12.5 \si{\angstrom}, \emph{c}=7.70 \si{\angstrom}} Further analysis was therefore required to confirm the exact phase of \ce{NaOH} \ce{WO3} NWs.\cite{Azimirad2009a}

\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{naoh_wo3_tem.jpg}
\caption[Characterization of \ce{NaOH} \ce{WO3}: TEM]{TEM Characterization of \ce{NaOH} \ce{WO3}: (a) TEM image of one nanowire with diameter ca. 20 nm, and (b) HRTEM images on the tip region showing absence of catalytic particle.}
\label{fig:naohwtem}
\end{figure}

TEM specimen was prepared by using carbon grid to slightly scratch the as-grown sample. Fig.~\ref{fig:naohwtem} show the feature of majority NWs. The growth direction was difficult to determined due to the presence of streaking in SAED pattern (not shown here), which presumably arose from stacking defaults during the NW growth. HRTEM revealed dominant plane spacings of 3.90 and 3.80 \si{\angstrom}, in favorable comparison to the XRD peak at $22.92^\circ$ and $23.66^\circ$. It was worth noting that catalyst particles were absent on the tip of NWs during the TEM observation. 
\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{Na2WO4_phase}
\caption[Phase diagram of \ce{Na2WO4}-\ce{WO3}]{Phase diagram of \ce{Na2WO4}-\ce{WO3} reproduced from Ref.~\cite{Hoermann1929}}
\label{fig:nawopd}
\end{figure}
Based on the above characterizations, this thesis concluded that the majority of NW belonged to tungsten oxide, although Na doping cannot be ruled out completely. According to the \ce{Na2WO4}-\ce{WO3} phase diagram in Fig.~\ref{fig:nawopd}, the eutectic temperature is ca. 720 \si{\degreeCelsius}.\cite{Mann2007} Further investigations are required to better understanding the growth mechanism. The author suggested using other alkali metal compound as catalyst to probe the \ce{WO3} growth.

\section{Preliminary results of MoO$_3$ growth on glass}\label{sec:mogls}

The various growth morphologies found using alkali metal based materials as catalysts demonstrated the proposed \gls{vss} mechanism and transverse mode was a general phenomenon, and broad engineering space could be explored to control the dimension of \ce{MoO3}. As an example, this study further illustrated the tunable growth of \ce{MoO3} nanoribbons on glass substrate. Achieving morphology controlled growth of \ce{MoO3} on transparent substrates, e.g. glass, is an important progress towards process integration and device fabrication in photon-energy related applications.

This thesis studied the effect of temperature, oxygen partial pressure, and growth time towards \ce{MoO3} deposition on glass substrates. Each factor was divided into two levels with $T$ at 700 and 800 \si{\degreeCelsius}, \ce{O2} flow at 1 and 10 sccm, and growth time of 15 and 60 min, respectively. And the morphological variations revealed by SEM imaging were summarized in Table~\ref{tab:mo3glass}.
% glass growth matrix
\begin{table}[htb]
\centering
\caption{\ce{MoO3} growth on glass}\label{tab:mo3glass}
\begin{tabular}{lccp{3in}}
\toprule
\multicolumn{3}{c}{Growth Conditions} \\
\cmidrule(l){1-3}
$T$ (\si{\degreeCelsius}) & \ce{O2} (sccm) & Time (min) & Morphology  \\
\midrule
700    &  10   & 15  &   about 50 $\mu$m belts, some towers \\
700    &  1    & 60  &   irregular shape, not too much growth\\
700    &  10   & 60  &   particle almost all on tip of belts, many towers \\
800    &  1    & 15  &   tens $\mu$m belt with sharp edge(100 nm thick, and 1micro wide) dominates\\
800    &  1    & 60  &   exhibiting ITO features, particles on tips of belt, many tower like structures\\
800    &  10   & 15  &   rounded edge flakes dominate, modest density tapering belt, and some possible tower\\
800    &  10   & 60  &   ITO/glass features, aggregated belts, some towers\\
\bottomrule
\end{tabular}
\end{table}
The 800-1-15 combination produced dense \ce{MoO3} nanoplate with length about 10 $\mu$m, width $300\sim500$ nm and thickness less than 100 nm, as illustrated in Fig.~\ref{fig:ch4glass}. These \ce{MoO3} nanoplates were of high crystalline quality and could potentially show enhanced photoelectrochemical and electrochromic activities. 
%010714r1 sample
\begin{figure}[htb]
\centering
\includegraphics[width=0.7\textwidth]{moo3_glass_lbd}
\caption[Optimized \ce{MoO3} nanoplates growth on glass]{Optimized \ce{MoO3} nanoplates growth on glass.}
\label{fig:ch4glass}
\end{figure}
%raman data on moo3 glass available
For future work, the author suggested several following perspectives:
\begin{enumerate}
\item further investigation of \ce{WO3} growth using alkali metal ions as catalysts;
\item probing the VSS growth method of \ce{MoO3} using HRTEM and analyzing the catalyst state; and
\item applying TEM-Raman integrated study onto other nanostructures.
\end{enumerate}