### 2015-03-29 main.tex

\documentclass[12pt,oneside]{book}
   \usepackage{unccx} % uncc format specification seems this style should be loaded before other packages
   \input{package_iff.tex} % additonal packages and tweaking
   \input{./FrontMatter/abbre}

\begin{document}
    \input{./FrontMatter/FrontMatter}
   \include{./ch-intro/ch-intro}
   \include{./ch-intro/ch-instr}
   \include{./ch-wox/ch-wox}
   \include{./ch-moox/moox} % no tex needed, otherwise there is error
   \include{./ch-tms/tms}

%\appendix
  % \glsaddallunused
   %\include{./BackMatter/app1}
   %\include{./BackMatter/app2}
   %\include{./BackMatter/app_training}
   % \include{./BackMatter/app-tms}
%   \include{./BackMatter/paper}

\renewcommand{\bibname}{REFERENCES}
\begin{singlespace}
  \printbibliography
\end{singlespace}

\end{document}

Similar phenomena were observed in other studies. \citeauthor{Thangala2007} reported that a decrease in NW density with increasing substrate temperature, and an increase of NW density with increasing partial pressure of oxygen.\cite{Thangala2007}

When the catalyst was smaller than some critical size, there is no preferred growth direction. This changed when the precipitated solid exceeded certain critical dimension, and the morphology evolution was then determined by the relative size of already formed solid and the associated catalyst.

forming a swelling particle

 the tips of \ce{WO3} NWs exhibited greater crystalline disorder, and \ce{WS2} layers on tips were folded at least along two axes, resulting a higher density of defects and thus a higher sulfurization rate; while the folding along NW body was quasi-one dimensional, subsequently the sulfurization rate could be lower. Besides, sulfurization could also be initiated from the NW body, presumably due to the presence of surface defects. As a result, this led to the uneven \ce{WS2} layer number distribution along the core-shell NWs.

which was tentatively attributed to stacking fault in this study.

should check how other template deals with before after section skip
 the belowcaptionskip is \the\belowcaptionskip, and the above is \the\abovecaptionskip, the intextsep is \the\intextsep. and the baselineskip is \the\baselineskip.